### 2014-01-16 docs/paper/paper.tex
paper outline and working ethernet transmit path, with modified integrator example

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{babel}
\begin{document}

\section{Introduction}

\subsection{Motivation}

\subsection{Potential Applications}

\subsection{State of the Art (Neuromorphic Hardware)}

\section{Methods}

\subsection{Description of NEF and population mode}

\subsubsection{Considerations related to hardware (multiplexing, clustering, using principal components)}

\subsection{High-level system description}

\subsubsection{Time multiplexing}

\subsection{Hardware Design details}

\subsubsection{pipelining}

\subsubsection{each component}

\subsection{Constraints on models}

\subsection{Implementation details}

\subsubsection{VC707}

\subsubsection{Ethernet}

\subsection{Nengo backend}

% quick description of what that involves

\section{Results}

\subsection{basic simulation}

\subsection{elaborate simulations}

\subsection{power, \% timestep used for each simulation, \% of board resources for each simulation}

\section{Discussion}

\subsection{How well it worked}

\subsection{Limitations}
\subsection{Future work \& extensions}

\end{document}

### 2014-01-29 docs/paper/paper.tex
completed hardware design section

@@
 \usepackage{babel}
+
+\newcommand{\design}{Nengo-RT}
+
 \begin{document}
@@
 
-\subsubsection{Time multiplexing}
+\design{} is a custom digital hardware design that implements the population-mode NEF in real-time (1~ms timestep).
+% FIXME figure: block diagram
+The encoder units perform the function of encoding (), % FIXME NEF equation
+while the population units perform the decoding () % FIXME NEF equation
+and write the decoded values to the decoded value buffers
+to be used as inputs to other populations or as outputs from the network.
+Each population of neurons is assigned to a population unit, which is able to simulate up to 1024 populations
+having either one or two encoded inputs each. These inputs are calculated by the encoder units,
+of which there are two for a one-dimensional population unit and four for a two-dimensional population unit.
+Within each population unit, the decoder weights for each specific population are time-multiplexed
+in order to run more populations on the same hardware, while the principal components are fixed and must be shared
+between all populations on an individual population unit.
+Each population unit writes to a fixed set of decoded value buffers, but a shared interconnect
+allows the encoder units to perform random accesses to these buffers, which allows
+inputs to a population to come from any combination of decoded values.
+A number of decoded value buffers are reserved for external inputs to the network,
+which can be sent from the controlling computer system over an Ethernet connection.
+The output channels can also be programmed to send decoded values back to the computer over Ethernet.
 
-\subsection{Hardware Design details}
+\subsection{Hardware design details}
 
-\subsubsection{pipelining}
+\subsubsection{Encoder Unit}
 
-\subsubsection{each component}
+Each encoder unit communicates with the decoded value buffers over a shared interconnect, described below.
+This is done in order to allow every encoder to read from every decoded value buffer.
+However, because the decoded value buffers have a fixed number of access ports for reading data,
+a maximum of two encoder units are allowed to access a single decoded value buffer simultaneously
+without causing a bus collision.
+
+Each encoder unit is attached to a circular buffer of size 8192
+containing a list of instructions for calculating one encoded value. 
+This size was chosen as a compromise between memory utilization and encoder flexibility.
+Each instruction corresponds to delaying for one or more clock cycles
+(in order to coordinate concurrent access to the interconnect),
+followed by reading one decoded value a certain address.
+The decoded value data is made available to the encoder unit three clock cycles later
+(due to interconnect delays), at which time it is multiplied by an encoder weight
+and added to the accumulated encoded value, initially 0.
+A certain bit in the instruction field acts as a flag to indicate that the encoded value is ready to be passed on to the population unit after this
+instruction finishes executing. The encoder unit then waits for all other encoders to finish in order to synchronize
+access to the shared interconnect, then resets the accumulator and begins encoding the next value.
+
+The encoder units must each perform 1024 encoding operations per timestep.
+To avoid duplication of logic, a single encoding controller connects to all encoder units and is responsible for
+signalling them to start encoding when it detects that they have all finished encoding the previous value.
+The encoding controller counts how many values have been encoded per timestep and
+ensures that after 1024 such operations, the encoder units are not signalled to start again until
+the beginning of the next timestep.
+
+\subsubsection{Population Unit}
+
+Encoded values are passed from the encoder units to the population units. Population units are either one-dimensional
+(functions of one variable) or two-dimensional (functions of two variables).
+Both types of population unit receive input from two encoder units in each input dimension;
+for each dimension, the encoded values are passed through a separate first-order filter to simulate % FIXME ...post-synaptic time constant?
+and then summed. The value resulting from this operation is
+truncated to the most significant 10 bits and is then used as input to a set of lookup tables,
+implemented as a bank of synchronous RAM elements. SRAMs are used
+because the contents of these tables are not known until simulation-time and therefore the lookups cannot be implemented as hardware logic functions.
+Each lookup table contains the 12-bit fixed-point values of points sampled from the principal components shared by all populations of neurons
+simulated on that population unit. For one-dimensional population units, there are seven such lookup tables, and each one contains
+1024 values sampled at equally-spaced points on the open interval $(-2, 2)$. For two-dimensional population units,
+there are fifteen lookup tables, each containing 1024 values sampled at equally-spaced points on a square of width 4
+centered at $(0, 0)$. The resulting sample space is a 32x32 grid.
+To increase the accuracy of this representation, each two-dimensional population unit performs bilinear interpolation
+using the least significant five bits of each input variable to determine the distance between sample points. 
+The extra resources used to perform this operation are modest
+and significantly reduce the amount of memory used for two-dimensional populations while maintaining a good degree of accuracy of representation.
+
+To simulate the noise resulting from a neural representation, a pseudorandom Gaussian noise generator is implemented in each population unit.
+The output from the noise generator is passed through a programmable second-order filter to approximate the power spectrum of the noise
+that would be present if individual neurons were being implemented. % FIXME fact check and possibly get a citation
+
+The interpolated values from the principal component lookup tables and the filtered noise sample are multiplied by a vector of decoders,
+and these results are summed to produce a decoded value. Each population unit can store up to four independent sets of decoders per population being simulated,
+which allows each population to produce a maximum of four decoded values as some function of its inputs
+(and the resulting values of its principal components).
+
+\subsubsection{Decoded Value Buffers and Interconnect}
+
+Decoded values calculated from population units are stored in decoded value buffers, which are fairly large memory banks that can each
+store 2048 decoded values. The buffers themselves expose a read-only interface, accessed by the encoders to read decoded values from the
+previous timestep, and a write-only interface, accessed by the population units when new decoded values are calculated.
+Internally, each interface is connected to a separate dual-port RAM; the connections are swapped between RAMs
+at the end of each timestep in order to implement double-buffering, so that it is not possible for encoders to read a
+partially-updated set of decoded values, where some are from the previous timestep and some are freshly written during the current timestep.
+% FIXME DV double buffer diagram in here somewhere
+Because each population unit produces 4096 decoded values (1024 DVs from each of four decoder vectors),
+this means that a population unit must be attached to the write-only interfaces of two separate decoded value buffers.
+
+The encoders must share access to the read-only interface of the decoded value buffers. To accomplish this,
+the encoders are connected to the decoded value buffers over a shared interconnect. Access to the decoded value buffers
+is managed by arbiters that inspect the decoded values being read from by encoders and redirect the read operations to the
+corresponding buffers. A maximum of two encoders may be accessing a single buffer at any one time.
+Therefore, it is up to the compiler to generate a sequence of read instructions for the encoders that ensures that this
+constraint is respected. The delay field in the encoder instruction word can be used for this purpose to make an encoder
+wait until another is finished with its access before attempting to read from the same buffer.
+
+\subsubsection{Input, Output, and Control}
+
+A number of decoded value buffers are set aside for use as inputs to the simulation. These are indistinguishable from other decoded value buffers
+from the perspective of the encoders and interconnect, but instead of being written to by population units,
+they are connected to one or more input channels which are responsible for receiving data from an external source
+and writing it to the buffers.
+
+To provide output to an external source, such as a simulation GUI running on a computer, one or more output channels
+can be implemented whose function is to read from the decoded value buffers and output the desired decoded values to be transmitted
+to an external device. Output channels do not have access to the interconnect until all encoders have finished all
+encoding operations for this timestep. At that point, the encoders are detached from the interconnect by a separate controller and the output
+channels are attached in their place. This allows the output channels to share interconnect logic with the encoders;
+it also increases efficiency both by reducing contention on the shared interconnect (as the encoders and output channels
+do not access it at the same time) and by allowing output to happen at the same time as decoding
+(the double-buffering of decoded values means that output channels always see a consistent set of values,
+so these operations can be performed in parallel).
+
+Both the input and output channels are transport-agnostic and can be attached to application-specific custom hardware that allows for
+I/O to be performed over the desired interface, such as Ethernet.
+
+The top-level module exposes a programming interface that allows for any of the runtime configurable components
+(decoded value buffers, encoders, population unit filters, principal component lookup tables,
+population unit noise generators, decoder vectors, and I/O channels) to be initialized and programmed
+over an external interface such as Ethernet. This allows a single implementation of the design
+to be used to simulate a large number of different neural networks without requiring modifications to the physical hardware.
+The programming interface also allows an external controller to reset the hardware and to start, pause, or single-step a simulation
+without requiring physical access to the hardware.
 

### 2014-01-29 docs/paper/paper.tex
point form notes under certain other sections

@@
 
+\begin{itemize}
+\item encoder instruction buffer size limits number of decoded values that can be encoded
+\item dimensional restrictions -- 1D, 2D population units
+\item four decoded values per population -- using more will create copies of populations to accommodate extra decoders
+\end{itemize}
+
 \subsection{Implementation details}
 
-\subsubsection{VC707}
-
-\subsubsection{Ethernet}
+\begin{itemize}
+\item VC707 board
+\item resources available -- BRAM, DSP
+\item implemented design -- number of population units, DV buffers
+\item FPGA resources utilized; maximum theoretical design size
+\item clock speed and implications of this
+\item Ethernet communication for control and I/O -- maximum channel capacity in terms of DVs
+\end{itemize}
 
@@
 
-% quick description of what that involves
+\begin{itemize}
+\item quick description of Nengo
+\item Nengo ensembles are hardware populations
+\item Nengo connections are hardware encoders and decoders
+\item Nengo nodes and probes are hardware I/O endpoints
+\item compiler functions: calculate principal components, map populations to PCs, calculate decoders and encoders,
+calculate encoder schedule, create programming file
+\item Nengo will then program the board and communicate with it during simulation
+\end{itemize}
 
@@
 \subsection{Limitations}
+
 \subsection{Future work \& extensions}
 
+\begin{itemize}
+\item board-to-board communication for larger networks
+\item increase design size by using off-chip memory (DDR3)
+\item better network stack to allow control and I/O over common protocols e.g. UDP
+\end{itemize}
+
 \end{document}

### 2014-01-30 docs/paper/paper.tex
constraints on models

@@
 
-Each encoder unit is attached to a circular buffer of size 8192
+Each encoder unit is attached to a circular buffer of size 8192 % FIXME check the size
 containing a list of instructions for calculating one encoded value. 
@@
 \item encoder instruction buffer size limits number of decoded values that can be encoded
+
+There is an artificial limitation on the complexity of the encoding operation due to the length of the instruction buffer that is used.
+With a size of 8192, % FIXME check size
+because each encoder must produce 1024 encoded values per timestep, a maximum of eight instructions, and therefore eight decoded values
+(not necessarily distinct)
+can be used in encoding. It is theoretically possible to implement an encoding operation that requires more decoded values to be used,
+but the extra instructions used for this mean that other populations sharing that encoder will have to use fewer.
+It is expected that the compiler will be able to detect this scenario and reorganize populations appropriately so as to avoid overfilling the instruction buffers.
 \item dimensional restrictions -- 1D, 2D population units
+% FIXME citation about why 2D is ``good enough''
+
+As only one-dimensional and two-dimensional population units have been implemented, networks that require populations to compute functions
+of three or more independent variables cannot be simulated with this design.
 \item four decoded values per population -- using more will create copies of populations to accommodate extra decoders
+
+The choice of using a maximum of four sets of decoder vectors per population is arbitrary; it seems to be a good balance between
+population connectivity and memory usage. However, populations that have more than four outgoing connections must be treated specially.
+Assuming none of the outgoing connections can be combined together (for example, if the function they decode only differs by a constant),
+in order to implement such a population on the simulator, it is necessary to duplicate it and simulate it twice,
+once with the first four decoder vectors and again with the remaining ones. The consequence of this is that a population with more than
+four outgoing connections will be represented as two or more populations in simulation, thereby reducing the actual number of
+distinct populations that can be simulated.
 \end{itemize}

### 2014-01-30 docs/paper/paper.tex
future work, partial

@@
 \item board-to-board communication for larger networks
+
+The flexibility of input and output channels lends to their use in connecting multiple devices together for use as part of a larger simulation engine.
+This would allow networks to be simulated that consume more resources than are available on a single device, which may in turn allow
+for ``grids'' of inexpensive hardware devices to be used to simulate a very large model.
 \item increase design size by using off-chip memory (DDR3)
 \item better network stack to allow control and I/O over common protocols e.g. UDP
+
+The current control and I/O interface communicates over Ethernet, but since bare frames are being used with no higher-level protocols on top of them,
+it is somewhat cumbersome to set up a computer or other device to send and receive data to the hardware.
+This approach is suitable for testing and prototyping, but it makes certain assumptions about the setup of the network and the host computer that
+may not be reasonable for general use, such as requiring raw socket privileges on the host operating system (which typically require administrator access).
+A more user-friendly approach would be to implement an Ethernet stack that supports a protocol such as UDP, which can be readily used by many different
+kinds of host software in order to interact with a simulation.
 \end{itemize}

### 2014-03-17 docs/paper/paper.tex
preliminary work on results section of paper

@@
 \usepackage{babel}
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{listings}
 
@@
 \begin{document}
+
+\lstset{basicstyle=\footnotesize}
 
@@
 
-\subsection{basic simulation}
+To verify the correct operation of the hardware as well as the software backend to Nengo, a simple network consisting of a single neural integrator was compiled and simulated on the hardware.
+Figure~\ref{lst:integrator1d} shows the code listing and the resulting network that it describes.
+\begin{figure}
+\centering
+
+\begin{subfigure}[b]{\textwidth}
+\centering
+\lstset{language=Python}
+\begin{lstlisting}[frame=single]
+import nengo
+import nengo.helpers
+model = nengo.Model(label='Integrator')
+input = nengo.Node(nengo.helpers.piecewise({0:0.5}), label='Input')
+tau = 0.1
+A = nengo.Ensemble(nengo.LIF(100, tau_rc=0.02, tau_ref=0.002,
+                   dimensions=1, label='Integrator'))
+nengo.Connection(A, A, transform=[[1]], filter=tau)
+nengo.Connection(input, A, transform=[[tau]], filter=tau)
+p1 = nengo.Probe(input, 'output')
+p2 = nengo.Probe(A, 'decoded_output', filter=0.01)
+\end{lstlisting}
+\caption{Source code}
+\label{lst:integrator1d:code}
+\end{subfigure}
+
+\begin{subfigure}[b]{0.3\textwidth}
+\centering
+\includegraphics[width=3.0in]{integrator-1d-schematic.eps}
+\caption{Schematic}
+\label{lst:integrator1d:schematic}
+\end{subfigure}
+
+\caption[A 1D neural integrator in Nengo.]
+{Source code (\ref{lst:integrator1d:code}) and schematic (\ref{lst:integrator1d:schematic}) of a 1D neural integrator in Nengo.}
+\label{lst:integrator1d}
+\end{figure}
+
+% FIXME explanation of the code?
+The network was compiled with the Nengo-RT backend to produce a programming file that could be used to configure the simulator hardware.
+Simulation output from the FPGA was captured with an Ethernet traffic sniffer and plotted in Figure~\ref{fig:integrator1d}.
+
+\begin{figure}
+\centering
+
+\includegraphics[width=6in]{integrator-1d.eps}
+
+\caption[Simulation of a 1D neural integrator.]
+{Hardware output from simulation of a one-dimensional population representing a neural integrator.}
+
+\label{fig:integrator1d}
+\end{figure}
+
+% FIXME discussion of result?
 

### 2014-03-17 docs/paper/paper.tex
minor rework of limitations

@@
 Both the input and output channels are transport-agnostic and can be attached to application-specific custom hardware that allows for
-I/O to be performed over the desired interface, such as Ethernet.
+I/O to be performed over the desired interface, such as Ethernet. The implemented design performs I/O over Gigabit Ethernet using a custom Ethernet frame format for efficiency.
 
@@
 
-\begin{itemize}
-\item encoder instruction buffer size limits number of decoded values that can be encoded
-
 There is an artificial limitation on the complexity of the encoding operation due to the length of the instruction buffer that is used.
-With a size of 8192, % FIXME check size
-because each encoder must produce 1024 encoded values per timestep, a maximum of eight instructions, and therefore eight decoded values
-(not necessarily distinct)
-can be used in encoding. It is theoretically possible to implement an encoding operation that requires more decoded values to be used,
+Because each encoder must produce 1024 encoded values per timestep, using an instruction memory of size 8192 for each encoder means that a maximum of eight decoded values
+can be used as input to each encoder per population.
+It is theoretically possible to simulate populations that require more than eight decoded values to be read by a single encoder,
 but the extra instructions used for this mean that other populations sharing that encoder will have to use fewer.
 It is expected that the compiler will be able to detect this scenario and reorganize populations appropriately so as to avoid overfilling the instruction buffers.
-\item dimensional restrictions -- 1D, 2D population units
+
 % FIXME citation about why 2D is ``good enough''
-
 As only one-dimensional and two-dimensional population units have been implemented, networks that require populations to compute functions
 of three or more independent variables cannot be simulated with this design.
-\item four decoded values per population -- using more will create copies of populations to accommodate extra decoders
 
@@
 distinct populations that can be simulated.
-\end{itemize}
 
@@
 % FIXME explanation of the code?
-The network was compiled with the Nengo-RT backend to produce a programming file that could be used to configure the simulator hardware.
+The network was compiled with the \design{} backend to produce a programming file that could be used to configure the simulator hardware.
 Simulation output from the FPGA was captured with an Ethernet traffic sniffer and plotted in Figure~\ref{fig:integrator1d}.

### 2014-03-17 docs/paper/paper.tex
begin implementation section

--- 2014-01-30 docs/paper/paper.tex
@@
 
-Each encoder unit is attached to a circular buffer of size 8192 % FIXME check the size
+Each encoder unit is attached to a circular buffer of size 8192
 containing a list of instructions for calculating one encoded value. 
@@
 Both types of population unit receive input from two encoder units in each input dimension;
-for each dimension, the encoded values are passed through a separate first-order filter to simulate % FIXME ...post-synaptic time constant?
+for each dimension, the encoded values are passed through a first-order filter which performs post-synaptic filtering, % FIXME check this
 and then summed. The value resulting from this operation is
@@
 To simulate the noise resulting from a neural representation, a pseudorandom Gaussian noise generator is implemented in each population unit.
-The output from the noise generator is passed through a programmable second-order filter to approximate the power spectrum of the noise
+The output from the noise generator is passed through a pre-defined second-order filter to approximate the power spectrum of the noise
 that would be present if individual neurons were being implemented. % FIXME fact check and possibly get a citation
@@
 There is an artificial limitation on the complexity of the encoding operation due to the length of the instruction buffer that is used.
-With a size of 8192, % FIXME check size
+With a size of 8192,
 because each encoder must produce 1024 encoded values per timestep, a maximum of eight instructions, and therefore eight decoded values
@@
 of three or more independent variables cannot be simulated with this design.
+However, the majority of existing models in Nengo do not use populations with more than two input dimensions, so with the exception of
+certain very large or very complex models that require representation of high-dimensional quantities, providing only one- and two-dimensional
+populations is not expected to be overly restrictive.
+
 \item four decoded values per population -- using more will create copies of populations to accommodate extra decoders
@@
 \item resources available -- BRAM, DSP
+
+A hardware implementation of the Nengo-RT design was produced on a Xilinx VC707 development board, which includes a modern field-programmable gate array (FPGA)
+device (Xilinx XC7VX485T-2FFG1761C). This device was chosen because of its ready availability on a development platform and its large logic capacity and available resources.
+In particular, it provides 37~080~Kb of on-chip block RAM and 2~800 customizable DSP slices, both of which the Nengo-RT design requires in abundance in order to support
+the largest possible implementation size.
+
 \item implemented design -- number of population units, DV buffers
+
 \item FPGA resources utilized; maximum theoretical design size
@@
 \item Ethernet communication for control and I/O -- maximum channel capacity in terms of DVs
+
+The VC707 development board provides a Gigabit Ethernet PHY that can be used for fast off-board communication to a PC.
+A simple Ethernet transceiver was implemented as part of the Nengo-RT design to expose a programming interface to the Nengo software,
+as well as to provide input and output of decoded values during simulation.
+
 \end{itemize}

### 2014-03-17 docs/paper/paper.tex

--- 2014-03-17 docs/paper/paper.tex
@@
 of three or more independent variables cannot be simulated with this design.
+However, the majority of existing models in Nengo do not use populations with more than two input dimensions, so with the exception of
+certain very large or very complex models that require representation of high-dimensional quantities, providing only one- and two-dimensional
+populations is not expected to be overly restrictive.
 

### 2014-03-17 docs/paper/paper.tex
clock implications

@@
 
-A hardware implementation of the Nengo-RT design was produced on a Xilinx VC707 development board, which includes a modern field-programmable gate array (FPGA)
+A hardware implementation of the \design{} design was produced on a Xilinx VC707 development board, which includes a modern field-programmable gate array (FPGA)
 device (Xilinx XC7VX485T-2FFG1761C). This device was chosen because of its ready availability on a development platform and its large logic capacity and available resources.
@@
 \item clock speed and implications of this
+
+The implemented hardware operates at a clock speed of 125~MHz, using the same extermal clock source as the Gigabit Ethernet PHY.
+This is convenient as it means that the entire design, including both the simulator hardware and the I/O interfaces, operates from the same clock source.
+Because of this, there is no latency when receiving input or transmitting output to and from the simulation;
+if two different clock speeds were used, extra buffers would have to be implemented in order to ``cross'' the two clock domains,
+and the logic operating in the faster clock domain would occasionally have to wait for the slower-clocked logic to consume input from the buffer.
+This type of stall would prevent part of the simulator from operating in real-time; by using the same clock for all logic, however, this problem is averted and
+both input and output are wait-free.
+
 \item Ethernet communication for control and I/O -- maximum channel capacity in terms of DVs

### 2014-04-02 docs/paper/paper.tex
backend description and 2D oscillator

@@
 
-\begin{itemize}
-\item VC707 board
-\item resources available -- BRAM, DSP
-
 A hardware implementation of the \design{} design was produced on a Xilinx VC707 development board, which includes a modern field-programmable gate array (FPGA)
@@
 
-\item implemented design -- number of population units, DV buffers
-
-\item FPGA resources utilized; maximum theoretical design size
-\item clock speed and implications of this
+% FIXME implemented design size
 
@@
 
-\item Ethernet communication for control and I/O -- maximum channel capacity in terms of DVs
-
 The VC707 development board provides a Gigabit Ethernet PHY that can be used for fast off-board communication to a PC.
@@
 
-\end{itemize}
+% FIXME maximum channel capacity in terms of decoded values
 
@@
 
+% FIXME quick description of Nengo
+Nengo's ensembles of neurons become populations that are simulated in the hardware on population units.
+Connections between populations are handled through encoders and decoders.
+Nodes and probes, which collect data or evaluate arbitrary functions in software, are
+simulated on the host computer and are transferred to and from the board over hardware 
+I/O channels.
+
+A backend to Nengo performs the task of compiling a Nengo model to a loadfile that can be
+simulated on the hardware. The compiler calculates principal components for each population
+of neurons, and then performs a clustering operation in order to associate many
+populations with each set of principal components while minimizing the differences
+between the approximate principal components and the actual values.
+The compiler calculates decoders and encoders for each population with respect
+to the shared principal components, and computes an optimized encoder schedule
+that coordinates reads made by each encoder so that all operations can be performed
+in as few clock cycles as possible while avoiding a read collision on any decoded value block.
+Once the compiler has finished writing out the loadfile, the Nengo simulator
+transfers the configuration data to the hardware and controls input and output during the simulation.
+
 \section{Results}
+
+% integrator (1d)
+% oscillator (2d)
+% feed-forward nonlinear functions
+% 2D multiplication
+
+% scope of situations for which it works well
+% full neurons vs. CPU populations vs. hardware populations
+% high-frequency breakdown (square wave, sinusoids)
+% qualitative comparison: show the figure, talk about the figure
+% possible quantitative approach (RMSE)
+
+% hardware-specific results: how many populations, 
+% circumstances under which an overflow could occur
+% (probabilistic analysis), power analysis,
+% resource utilization
 
@@
 
-% FIXME discussion of result?
+Given a constant input value of 0.5, the output from the integrator approaches this value
+over time and reaches it after 1.0 seconds, then continues to accumulate the input until
+another second has elapsed. The population saturates and remains at 1.0 after this time
+because it is not able to represent values with a magnitude greater than 1.0. However,
+the integrator's behaviour for the first two seconds of simulation is correct and confirms
+the proper operation of the one-dimensional population hardware.
 
-\subsection{elaborate simulations}
+The two-dimensional population unit was tested with a simple single-population van der Pol oscillator.
+Figure~\ref{lst:oscillator2d} shows the code listing and the network generated in Nengo;
+Figure~\ref{fig:oscillator2d} shows output from the \design{} hardware simulation.
+\begin{figure}
+\centering
+
+\begin{subfigure}[b]{\textwidth}
+\centering
+\lstset{language=Python}
+\begin{lstlisting}[frame=single]
+import nengo
+import nengo.helpers
+model = nengo.Model('Oscillator')
+neurons = nengo.Ensemble(nengo.LIF(200), dimensions=2, label='van der Pol Oscillator')
+input = nengo.Node(output=nengo.helpers.piecewise({0:[1,0], 0.1:[0,0], label='Initial condition')
+tau = 0.1
+freq = 5.0
+input2output = nengo.Connection(input, neurons)
+recurrent = nengo.Connection(neurons, neurons, transform=[[1, -freq*tau],[freq*tau,1]], filter=tau)
+input_probe = nengo.Probe(input, 'output')
+neuron_probe = nengo.Probe(neurons, 'decoded_output', filter=0.005)
+\end{lstlisting}
+\caption{Source Code}
+\label{lst:oscillator2d:code}
+\end{subfigure}
+
+\begin{subfigure}[b]{0.3\textwidth}
+\centering
+% FIXME \includegraphics[width=3.0in]{oscillator-2d-schematic.eps}
+\caption{Schematic}
+\label{lst:oscillator2d:schematic}
+\end{subfigure}
+
+\caption[A 2D neural oscillator in Nengo.]
+{Source code (\ref{lst:oscillator2d:code}) and schematic (\ref{lst:oscillator2d:schematic})
+of a 2D neural oscillator in Nengo.}
+\label{lst:oscillator2d}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=6in]{oscillator-2d.eps}
+
+\caption[Simulation of a 2D neural oscillator.]
+{Hardware output from the simulation of a two-dimensional population representing a neural oscillator.}
+\label{fig:oscillator2d}
+\end{figure}
+
+The oscillator requires a brief non-zero ``start-up'' input in order to begin oscillating.
+After this input disappears (0.1~s), both dimensions of the decoded output can be seen to
+oscillate in a sinusoidal fashion. This oscillation, once started, will proceed indefinitely.
+% FIXME frequency analysis?
+The oscillation as observed in the figure confirms the correct operation of the two-dimensional
+population hardware.
 

### 2014-04-03 docs/paper/paper.tex
fix overset listing for oscillator

@@
 model = nengo.Model('Oscillator')
-neurons = nengo.Ensemble(nengo.LIF(200), dimensions=2, label='van der Pol Oscillator')
-input = nengo.Node(output=nengo.helpers.piecewise({0:[1,0], 0.1:[0,0], label='Initial condition')
+neurons = nengo.Ensemble(nengo.LIF(200), dimensions=2, 
+  label='van der Pol Oscillator')
+input = nengo.Node(output=nengo.helpers.piecewise(
+  {0:[1,0], 0.1:[0,0], label='Initial condition')
 tau = 0.1
@@
 input2output = nengo.Connection(input, neurons)
-recurrent = nengo.Connection(neurons, neurons, transform=[[1, -freq*tau],[freq*tau,1]], filter=tau)
+recurrent = nengo.Connection(neurons, neurons, 
+  transform=[[1, -freq*tau],[freq*tau,1]], filter=tau)
 input_probe = nengo.Probe(input, 'output')
@@
 
+% things that Nengo can do that we can't do:
+% >2d representation
+% inhibitory connections
+% learning
+% high-frequency stuff (this will depend on other results)
+
 \subsection{Future work \& extensions}

### 2014-04-11 dump2val.py
modifications for multiplier example

@@ -27,7 +27,12 @@ for line in open(sys.argv[1]):
         line = line[3:]
         i += 1
 
+ch = 0
 for pts in dataPoints:
-    plt.plot(pts)
+    plt.plot(pts, label="Channel " + str(ch))
+    ch += 1
 
+print(str(ch) + " channels")
+
+plt.legend(loc='best')
 plt.show()

### 2014-04-11 programming_test.py

@@ -4,10 +4,40 @@ from socket import socket, AF_PACKET, SOCK_RAW
 from time import sleep
 import sys
 import struct
+import math
 
 src_addr = "\xf8\x1a\x67\x03\xa0\x72"
 dst_addr = "\x00\x0a\x35\x02\x8f\xc0"
 ethertype = "\x88\xb5"
+
+# Prepend p to the string s until its length is at least n.
+def pad(s, p, n):
+    x = s
+    while len(x) < n:
+        x = p + x
+    return x
+
+
+def float2sfixed(f):
+    if f < 0:
+        sgn = "1"
+        f = f * -1
+    else:
+        sgn = "0"
+    if f > 2047.0 / 1024.0:
+        f = 2047.0 / 1024.0 # largest representable value
+    ipart = int(math.floor(f))
+    fpart = int(round((f - ipart) * 2**10))
+    istr = pad(bin(ipart)[2:], "0", 1) # remove '0b' prefix and pad out
+    fstr = pad(bin(fpart)[2:], "0", 10)
+    # if the string is negative, take the 2's complement
+    if sgn == "1":
+        # take two's complement
+        n = int(istr + fstr, 2)
+        nc = 2**12 - n
+        return pad(bin(nc)[2:], "1", 12)
+    else:
+        return sgn + istr + fstr
 
 def makeframe(opcode, data):
     tag = "\x42" # FIXME better tag handling
@@ -55,13 +85,36 @@ for line in f:
     s.send(makeframe('\x00', binstring(addr) + binstring(data)))
     #sleep(0.001)
 
+sleep(1.0)
+
 print("Starting run...")
 s.send(makeframe_START())
 
-sleep(0.1)
+# DV input frame: |op = 0x01|addr|addr|pair count|data|data|data|...
+# write to (393216, 393217) (input bank #0 address #0 is board physical address #393216, etc.)
 
-print("Updating input DVs...")
-# write 393216 = 0.0, 393217 = 0.0 (input bank #0 address #0 is board physical address #393216, etc.)
-s.send(makeframe("\x01", "\x00\x00\x00\x00\x00\x00")) 
+# 0.0: (0.0, 1.0)
+# 1.5: (0.0, 0.8)
[…]

### 2014-04-14 docs/paper/paper.tex
initial multiplier example stuff

@@
 
+As a test of a combined network with both one-dimensional and two-dimensional populations,
+a neural multiplier was simulated on the hardware. Figure~\ref{lst:multiplier} shows
+the code listing and the resulting network that it describes.
+
+\begin{figure}
+\centering
+
+\begin{subfigure}[b]{\textwidth}
+\centering
+\lstset{language=Python}
+\begin{lstlisting}[frame=single]
+import nengo
+import nengo.helpers
+model = nengo.Model('Multiplier')
+A = nengo.Ensemble(nengo.LIF(100), dimensions=1,
+  radius=1.4, label='A')
+B = nengo.Ensemble(nengo.LIF(100), dimensions=1,
+  radius=1.4, label='B')
+combined = nengo.Ensemble(nengo.LIF(224), dimensions=2,
+  radius=2, label='combined')
+prod = nengo.Ensemble(nengo.LIF(100), dimensions=1,
+  radius=2, label='prod')
+
+inputA = nengo.Node(nengo.helpers.piecewise
+  ({0:0, 2.5:0.5, 4:-0.5}))
+inputB = nengo.Node(nengo.helpers.piecewise
+  ({0:1.0, 1.5:0.8, 3:0.0, 4.5:0.8}))
+
+nengo.Connection(inputA, A)
+nengo.Connection(inputB, B)
+
+nengo.Connection(A, combined, transform=[[1],[0]])
+nengo.Connection(B, combined, transform=[[0],[1]])
+
+def product(x):
+  return x[0] * x[1]
+
+nengo.Connection(combined, prod, function=product)
+
+A_probe = nengo.Probe(A, 'decoded_output', filter=0.01)
+B_probe = nengo.Probe(B, 'decoded_output', filter=0.01)
+prod_probe = nengo.Probe(prod, 'decoded_output', filter=0.01)
+\end{lstlisting}
+\caption{Source code}
+\label{lst:multiplier:code}
+\end{subfigure}
+
+\begin{subfigure}[b]{0.3\textwidth}
+\centering
+% \includegraphics[width=3.0in]{multiplier-schematic.eps} % FIXME
+\caption{Schematic}
+\label{lst:multiplier:schematic}
+\end{subfigure}
+
+\caption[A multiplier in Nengo.]
+{Source code(\ref{lst:multiplier:code}) and schematic (\ref{lst:multiplier:schematic})
+of a multiplier in Nengo.}
+\label{lst:multiplier}
+\end{figure}
+
+% FIXME: explain input and expected output
+
+Simulation of this network in hardware produced results comparable to those obtained in software
+(Figure~\ref{fig:multiplier}).
+
+\begin{figure}
+\centering
+
+% FIXME software subfigure
+
+\begin{subfigure}[b]{0.5\textwidth}
+\centering
+\includegraphics[width=3in]{multiplier.eps}
+\caption{Hardware}
+\label{fig:multiplier:hw}
+\end{subfigure}
+
+\caption[Multiplier simulation comparison.]
+{Comparison of software (FIXME) and hardware (\ref{fig:multiplier:hw})
+simulation of a neural multiplier network.}
+\label{fig:multiplier}
+\end{figure}
+
 \subsection{power, \% timestep used for each simulation, \% of board resources for each simulation}

### 2014-04-14 docs/paper/paper.tex
multiplier figures

@@
 \centering
-% \includegraphics[width=3.0in]{multiplier-schematic.eps} % FIXME
+\includegraphics[width=3.0in]{multiplier-schematic.eps}
 \caption{Schematic}
@@
 
-% FIXME software subfigure
+\begin{subfigure}[b]{0.5\textwidth}
+\centering
+\includegraphics[width=3in]{multiplier-sw.eps}
+\caption{Software}
+\label{fig:multiplier:sw}
+\end{subfigure}
 
@@
 \caption[Multiplier simulation comparison.]
-{Comparison of software (FIXME) and hardware (\ref{fig:multiplier:hw})
+{Comparison of software (\ref{fig:multiplier:sw}) and hardware (\ref{fig:multiplier:hw})
 simulation of a neural multiplier network.}

### 2014-04-22 docs/paper/paper.tex
first draft revised from comments

@@
 
+% FIXME brief description of FPGA hardware and programming
+
 \subsection{High-level system description}
@@
 having either one or two encoded inputs each. These inputs are calculated by the encoder units,
-of which there are two for a one-dimensional population unit and four for a two-dimensional population unit.
-Within each population unit, the decoder weights for each specific population are time-multiplexed
-in order to run more populations on the same hardware, while the principal components are fixed and must be shared
+of which there are two for a one-dimensional population unit and four for a two-dimensional population unit;
+each encoder unit can be associated with a different first-order filter to simulate post-synaptic filtering.
+For example, one of the two encoders on a one-dimensional population unit can be programmed with
+a small time constant to model fast synapses (e.g. AMPA and GABA$_A$), while the second can be programmed
+with a large time constant to model slow synapses (e.g. NMDA and GABA$_B$).
+Each population unit performs time-multiplexing and swaps in different sets of decoders
+in order to run more populations on the same hardware. The principal components are fixed and must be shared
 between all populations on an individual population unit.
@@
 allows the encoder units to perform random accesses to these buffers, which allows
-inputs to a population to come from any combination of decoded values.
+inputs to a population to come from any combination of decoded values. % FIXME more detail
 A number of decoded value buffers are reserved for external inputs to the network,
@@
 
+It is too resource-intensive to create a connection from the output of each population unit to the inputs of all population units (including itself).
+In order to allow communication of information across population units, the decoded values from each population are stored in a local buffer.
+Then, every population unit has a number of encoder units that are responsible for reading decoded values from each buffer. 
 Each encoder unit communicates with the decoded value buffers over a shared interconnect, described below.
-This is done in order to allow every encoder to read from every decoded value buffer.
-However, because the decoded value buffers have a fixed number of access ports for reading data,
-a maximum of two encoder units are allowed to access a single decoded value buffer simultaneously
-without causing a bus collision.
+A maximum of two encoder units are allowed to access a single decoded value buffer simultaneously
+without causing a bus collision, since the memory primitives used to implement the buffers have two independent read ports.
 
@@
 containing a list of instructions for calculating one encoded value. 
-This size was chosen as a compromise between memory utilization and encoder flexibility.
-Each instruction corresponds to delaying for one or more clock cycles
-(in order to coordinate concurrent access to the interconnect),
-followed by reading one decoded value a certain address.
-The decoded value data is made available to the encoder unit three clock cycles later
-(due to interconnect delays), at which time it is multiplied by an encoder weight
-and added to the accumulated encoded value, initially 0.
+This size was chosen as a compromise between memory utilization and encoder flexibility;
+it means that a single population can have a maximum fan-in of eight connections (including connections from the population to itself).
+To execute an instruction, the encoder sends the decoded value address it wants to read from to the interconnect
+and then reads the corresponding decoded value on its data input port three clock cycles later.
+The value is then multiplied by a fixed-point weight, which is given as part of the instruction field, and added to an accumulator in the encoder.
+The interconnect is responsible for routing address and value data between the encoders and the buffers.
+A portion of the instruction field provides a delay (expressed as a number of clock cycles) to the encoder before executing this instruction.
+This is necessary when many encoders are accessing the interconnect simultaneously in order to prevent too many of them from reading
+from the same buffer at the same time (which would result in data corruption, because only two concurrent accesses are possible).
 A certain bit in the instruction field acts as a flag to indicate that the encoded value is ready to be passed on to the population unit after this
@@
 access to the shared interconnect, then resets the accumulator and begins encoding the next value.
-
-The encoder units must each perform 1024 encoding operations per timestep.
-To avoid duplication of logic, a single encoding controller connects to all encoder units and is responsible for
+To synchronize encoders across population units, a single encoding controller connects to all encoder units and is responsible for
 signalling them to start encoding when it detects that they have all finished encoding the previous value.
-The encoding controller counts how many values have been encoded per timestep and
-ensures that after 1024 such operations, the encoder units are not signalled to start again until
-the beginning of the next timestep.
 
@@
 
-Encoded values are passed from the encoder units to the population units. Population units are either one-dimensional
-(functions of one variable) or two-dimensional (functions of two variables).
+Encoded values are passed from the encoder units to the population units. 
+Population units are either one-dimensional or two-dimensional.
+This means the groups of neurons they model have firing rates that are static functions of either a one- or two-dimensional represented variable.
 Both types of population unit receive input from two encoder units in each input dimension;
@@
 because the contents of these tables are not known until simulation-time and therefore the lookups cannot be implemented as hardware logic functions.
-Each lookup table contains the 12-bit fixed-point values of points sampled from the principal components shared by all populations of neurons
-simulated on that population unit. For one-dimensional population units, there are seven such lookup tables, and each one contains
+Each lookup table contains the 12-bit fixed-point values of points sampled from 
+%the principal components shared by all populations of neurons
+%simulated on that population unit. 
+principal components of the spike rates of neurons in a population, as a function of the variable that the population encodes.
+As described further below
+%(Section FIXME)
+, all populations on a population unit must share principal components.
+For one-dimensional population units, there are seven such lookup tables, and each one contains
 1024 values sampled at equally-spaced points on the open interval $(-2, 2)$. For two-dimensional population units,
@@
 centered at $(0, 0)$. The resulting sample space is a 32x32 grid.
+These numbers of principal components account for most of the variation in spike rates across 1D and 2D populations.
+% FIXME find out exactly how much of this variation they account for over random populations with a variety of parameters, and mention this.
 To increase the accuracy of this representation, each two-dimensional population unit performs bilinear interpolation
@@
 The output from the noise generator is passed through a pre-defined second-order filter to approximate the power spectrum of the noise
-that would be present if individual neurons were being implemented. % FIXME fact check and possibly get a citation
+that would be present if individual neurons were being implemented. % FIXME for completeness, give the transfer function
 
 The interpolated values from the principal component lookup tables and the filtered noise sample are multiplied by a vector of decoders,
-and these results are summed to produce a decoded value. Each population unit can store up to four independent sets of decoders per population being simulated,
+and these results are summed to produce a decoded value. % FIXME refer to equations
+One of the decoders corresponds to a gain term that can be used to control the amplitude of the noise.
+Each population unit can store up to four independent sets of decoders per population being simulated,
 which allows each population to produce a maximum of four decoded values as some function of its inputs
@@
 Decoded values calculated from population units are stored in decoded value buffers, which are fairly large memory banks that can each
-store 2048 decoded values. The buffers themselves expose a read-only interface, accessed by the encoders to read decoded values from the
+store 2048 decoded values. 
+Because each population unit produces 4096 decoded values (1024 DVs from each of four decoder vectors),
+this means that a population unit must be attached to the write ports of two separate decoded value buffers.
+The buffers themselves expose a read-only interface, accessed by the encoders to read decoded values from the
 previous timestep, and a write-only interface, accessed by the population units when new decoded values are calculated.
@@
 partially-updated set of decoded values, where some are from the previous timestep and some are freshly written during the current timestep.
+The reason for choosing a size of 2048 is because of this implementation detail; on the particular family of hardware
+that was used to implement the prototype, the largest single-component 12-bit dual-port RAM available has space for 1024 words,
+and two separate RAM components are necessary in order to implement the double buffer, hence a total size of 2048 words of memory.
 % FIXME DV double buffer diagram in here somewhere
-Because each population unit produces 4096 decoded values (1024 DVs from each of four decoder vectors),
-this means that a population unit must be attached to the write-only interfaces of two separate decoded value buffers.
 
@@
 the encoders are connected to the decoded value buffers over a shared interconnect. Access to the decoded value buffers
-is managed by arbiters that inspect the decoded values being read from by encoders and redirect the read operations to the
+is managed by arbiters that inspect the addresses requested by encoders and redirect the read operations to the
 corresponding buffers. A maximum of two encoders may be accessing a single buffer at any one time.
@@
 
-A number of decoded value buffers are set aside for use as inputs to the simulation. These are indistinguishable from other decoded value buffers
+A number of decoded value buffers, which can be customized at the time the hardware is synthesized
+(but always at least one)
+are set aside for use as inputs to the simulation. These are indistinguishable from other decoded value buffers
 from the perspective of the encoders and interconnect, but instead of being written to by population units,
@@
 
+% FIXME small diagram of hypothetical system in the loop (UI, robot, hardware, etc.)
+
 To provide output to an external source, such as a simulation GUI running on a computer, one or more output channels
-can be implemented whose function is to read from the decoded value buffers and output the desired decoded values to be transmitted
+are made accessible on the hardware. The exact number of output channels, as well as their features, will depend on the hardware used to implement the design.
+The function of the output channel is to read from the decoded value buffers and output the desired decoded values to be transmitted
 to an external device. Output channels do not have access to the interconnect until all encoders have finished all
@@
 
-There is an artificial limitation on the complexity of the encoding operation due to the length of the instruction buffer that is used.
+There is a limitation on the complexity of the encoding operation due to the length of the instruction buffer that is used.
 Because each encoder must produce 1024 encoded values per timestep, using an instruction memory of size 8192 for each encoder means that a maximum of eight decoded values
 can be used as input to each encoder per population.
-It is theoretically possible to simulate populations that require more than eight decoded values to be read by a single encoder,
+If a higher fan-in is required, the hardware design could be modified to allocate more memory to the instruction buffers,
+which in turn would reduce the amount of memory available for other portions of the design.
+It is theoretically possible to simulate populations that require more than eight decoded values to be read by a single encoder
+without increasing the size of the instruction buffers,
 but the extra instructions used for this mean that other populations sharing that encoder will have to use fewer.
-It is expected that the compiler will be able to detect this scenario and reorganize populations appropriately so as to avoid overfilling the instruction buffers.
+The compiler can detect this scenario and reorganize populations appropriately so as to avoid overfilling the instruction buffers.
 
@@
 of three or more independent variables cannot be simulated with this design.
-However, the majority of existing models in Nengo do not use populations with more than two input dimensions, so with the exception of
-certain very large or very complex models that require representation of high-dimensional quantities, providing only one- and two-dimensional
-populations is not expected to be overly restrictive.
+However, the majority of existing models in Nengo do not use populations with more than two input dimensions.
+A controlled oscillator is an example of a kind of NEF network that cannot be simulated in our hardware implementation,
+because it requires the approximation of a three-dimensional nonlinear function. Our implementation can simulate higher-dimensional representations
+by simulating dimensions individually or in pairs, as long as each approximated function is either linear or is nonlinear in only one or two dimensions.
 
 The choice of using a maximum of four sets of decoder vectors per population is arbitrary; it seems to be a good balance between
-population connectivity and memory usage. However, populations that have more than four outgoing connections must be treated specially.
+population connectivity and memory usage. Populations that have more than four outgoing connections must be treated specially.
 Assuming none of the outgoing connections can be combined together (for example, if the function they decode only differs by a constant),
-in order to implement such a population on the simulator, it is necessary to duplicate it and simulate it twice,
-once with the first four decoder vectors and again with the remaining ones. The consequence of this is that a population with more than
+in order to implement such a population on the simulator, it is necessary to duplicate it and simulate it
+$n/4$ times, once for each group of up to four decoder vectors.
+The consequence of this is that a population with more than
 four outgoing connections will be represented as two or more populations in simulation, thereby reducing the actual number of
@@
 device (Xilinx XC7VX485T-2FFG1761C). This device was chosen because of its ready availability on a development platform and its large logic capacity and available resources.
-In particular, it provides 37~080~Kb of on-chip block RAM and 2~800 customizable DSP slices, both of which the Nengo-RT design requires in abundance in order to support
+In particular, it provides 37080~Kb of on-chip block RAM and 2800 customizable DSP slices, both of which the Nengo-RT design requires in abundance in order to support
 the largest possible implementation size.
@@
 
-The implemented hardware operates at a clock speed of 125~MHz, using the same extermal clock source as the Gigabit Ethernet PHY.
+The implemented hardware operates at a clock speed of 125~MHz, using the same external clock source as the Gigabit Ethernet transceiver.
 This is convenient as it means that the entire design, including both the simulator hardware and the I/O interfaces, operates from the same clock source.
@@
 
-The VC707 development board provides a Gigabit Ethernet PHY that can be used for fast off-board communication to a PC.
-A simple Ethernet transceiver was implemented as part of the Nengo-RT design to expose a programming interface to the Nengo software,
+The VC707 development board provides a Gigabit Ethernet transceiver that can be used for fast off-board communication to a PC.
+A simple Ethernet stack was implemented as part of the Nengo-RT design to expose a programming interface to the Nengo software,
 as well as to provide input and output of decoded values during simulation.
@@
 
-\begin{itemize}
-\item quick description of Nengo
-\item Nengo ensembles are hardware populations
-\item Nengo connections are hardware encoders and decoders
-\item Nengo nodes and probes are hardware I/O endpoints
-\item compiler functions: calculate principal components, map populations to PCs, calculate decoders and encoders,
-calculate encoder schedule, create programming file
-\item Nengo will then program the board and communicate with it during simulation
-\end{itemize}
-
 % FIXME quick description of Nengo
@@
 simulated on the host computer and are transferred to and from the board over hardware 
-I/O channels.
+I/O channels after every timestep. 
+By default, the hardware operates in real-time; the software receives new output data from the board
+every millisecond and is responsible for processing this data and sending new inputs back to the hardware as quickly as possible.
+The software does not have to send inputs for every timestep; if no new data is sent, the old values will be preserved.
+This means that more work can be done in software while still operating the hardware in real-time by, for example,
+only computing new input values every fifth timestep.
+If strict synchronization is necessary, or if real-time operation is not possible,
+the hardware can be run in a single-step mode, which forces it to pause after completing each timestep in order
+to allow the software time to complete its tasks, which may include updating plots, logging received data to a file,
+computing node functions, and calculating and sending new input values to the hardware.
+This mode of operation, while not real-time, makes it possible to run hardware-assisted simulations
+on a slower PC or when the software component is too complex or slow to be performed in sync with the hardware
+(for example, logging a large number of decoded values to a hard drive).
 
@@
 populations with each set of principal components while minimizing the differences
-between the approximate principal components and the actual values.
+between the approximate principal components and the actual values. % FIXME explain/calculate why actual principal components cannot be used
 The compiler calculates decoders and encoders for each population with respect
-to the shared principal components, and computes an optimized encoder schedule
+to the shared principal components, and computes an optimized encoder schedule % FIXME details needed here
 that coordinates reads made by each encoder so that all operations can be performed
@@
 transfers the configuration data to the hardware and controls input and output during the simulation.
+
+% clustering algorithm, quantization error for large populations -- this needs its own section
 
@@
 % resource utilization
+
+% FIXME may need a simpler example first -- just an input driving an identity population from 0 to 0.5 in steps of 0.1. This is good for contrast.
 
@@
 
-\includegraphics[width=6in]{integrator-1d.eps}
+\includegraphics[width=6in]{integrator-1d.eps} % FIXME software plot
+% FIXME figure: larger axis and label font, make this consistent across all figures
 
@@
 Figure~\ref{fig:oscillator2d} shows output from the \design{} hardware simulation.
+
 \begin{figure}
@@
 
+% FIXME change the bounds on the axes to make it easier to see the similarities
+
 \begin{subfigure}[b]{0.5\textwidth}
@@
 
-% things that Nengo can do that we can't do:
-% >2d representation
-% inhibitory connections
-% learning
+Our current implementation does not support populations that represent variables of three or more dimensions.
+Support for higher-dimensional populations could be added in the future.
+However, as the number of dimensions increases, the
+amount of memory that must be used to store principal components and
+decoders increases as well; therefore, beyond a certain number of
+dimensions it would become impractical to implement a hardware
+population unit.
+
+Nengo supports a more advanced connection called a learning connection,
+which mdoifies the decoders of a population according to a particular ``learning rule''
+as a function of other represented values. The hardware design does not support this special connection type.
+It was decided to implement the most common cases first in order to demonstrate the feasibility of simulating
+neural networks in population mode on the hardware, and since many networks do not use learning connections,
+support for them was omitted from this version.
+
+Most connections in Nengo have both positive and negative synaptic weights. This
+structure is meant as a simplified model of a common excitatory/inhibitory circuit
+(Parisien et al.) % FIXME citation
+Nengo also has a special kind of connection called an inhibitory connection that
+reduces activity in all neurons of the target population. We did not model this kind of connection explicitly.
+However, such a connection onto a 1D population could be modelled by treating the population as 2D,
+setting the second element of each neuron's encoder to -1, and connecting inhibitory inputs to this dimension.
+
 % high-frequency stuff (this will depend on other results)
@@
 
-\begin{itemize}
-\item board-to-board communication for larger networks
-
 The flexibility of input and output channels lends to their use in connecting multiple devices together for use as part of a larger simulation engine.
@@
 for ``grids'' of inexpensive hardware devices to be used to simulate a very large model.
-\item increase design size by using off-chip memory (DDR3)
-\item better network stack to allow control and I/O over common protocols e.g. UDP
 
@@
 kinds of host software in order to interact with a simulation.
-\end{itemize}
+
+At present, the interconnect between the encoders and the decoded value
+buffers is the most complex part of the design in terms of the density of its logic
+in hardware. If there are $N$ encoders and $M$ decoded value buffers, a total of
+$NM$ connections must be made between these components in order for every
+encoder to be able to read from every decoded value buffer.
+Note that this is still a massive improvement over a neuron-to-neuron interconnect,
+which would require over 10000 times as many connections to provide all-to-all connectivity.
+% FIXME contrast this with size of neuron-to-neuron interconnect
+For large numbers
+of encoders and decoded value buffers, the complexity of the interconnect at
+that design size makes it impossible for the FPGA design tools to find a way to
+implement that part of the design on the device. This is currently the limiting
+factor in terms of the number of population units that can be implemented on
+the development board (as opposed to, for example, a lack of memory blocks for
+decoded value buffers). A future version of the \design{} design may address
+this issue and implement an interconnect that scales well to larger numbers of
+encoders and decoded value buffers in order to allow more populations to be
+simulated. % FIXME how could this be done?
 

### 2014-04-22 docs/paper/paper.tex
system block diagram
dv double buffer diagram

@@
 \design{} is a custom digital hardware design that implements the population-mode NEF in real-time (1~ms timestep).
-% FIXME figure: block diagram
+A block diagram of the design is shown in Figure~\ref{fig:system}.
 The encoder units perform the function of encoding (), % FIXME NEF equation
@@
 Each population unit writes to a fixed set of decoded value buffers, but a shared interconnect
-allows the encoder units to perform random accesses to these buffers, which allows
-inputs to a population to come from any combination of decoded values. % FIXME more detail
+allows the encoder units to perform random accesses to these buffers.
+This allows effective all-to-all communication between populations while reducing the number of 
+connections that must be made between population units.
 A number of decoded value buffers are reserved for external inputs to the network,
@@
 The output channels can also be programmed to send decoded values back to the computer over Ethernet.
+
+\begin{figure}
+\centering
+
+\includegraphics[width=6in]{system-block-diagram.eps}
+
+\caption{Block diagram of the \design{} hardware design.}
+
+\label{fig:system}
+\end{figure}
 
@@
 Both types of population unit receive input from two encoder units in each input dimension;
-for each dimension, the encoded values are passed through a first-order filter which performs post-synaptic filtering, % FIXME check this
-and then summed. The value resulting from this operation is
+for each dimension, the encoded values are passed through a first-order filter which performs post-synaptic filtering,
+and are then summed. The value resulting from this operation is
 truncated to the most significant 10 bits and is then used as input to a set of lookup tables,
@@
 Each lookup table contains the 12-bit fixed-point values of points sampled from 
-%the principal components shared by all populations of neurons
-%simulated on that population unit. 
 principal components of the spike rates of neurons in a population, as a function of the variable that the population encodes.
-As described further below
-%(Section FIXME)
-, all populations on a population unit must share principal components.
+In order to reduce the amount of memory used by each population, all populations on a population unit must share principal components.
 For one-dimensional population units, there are seven such lookup tables, and each one contains
@@
 partially-updated set of decoded values, where some are from the previous timestep and some are freshly written during the current timestep.
+A diagram of this component is shown in Figure~\ref{fig:dvdoublebuffer}.
 The reason for choosing a size of 2048 is because of this implementation detail; on the particular family of hardware
@@
 and two separate RAM components are necessary in order to implement the double buffer, hence a total size of 2048 words of memory.
-% FIXME DV double buffer diagram in here somewhere
+
+\begin{figure}
+\centering
+
+\includegraphics[width=6.0in]{dv-double-buffer-diagram.eps}
+
+\caption{Block diagram of the decoded value double-buffer component.}
+\label{fig:dvdoublebuffer}
+\end{figure}
 

### 2014-04-22 docs/paper/paper.tex
robot closed loop diagram

@@
 
-% FIXME small diagram of hypothetical system in the loop (UI, robot, hardware, etc.)
-
 To provide output to an external source, such as a simulation GUI running on a computer, one or more output channels
@@
 I/O to be performed over the desired interface, such as Ethernet. The implemented design performs I/O over Gigabit Ethernet using a custom Ethernet frame format for efficiency.
+Input and output channels could also be implemented for sensors and actuators to allow for autonomous operation of a Nengo network in hardware once it has been sent to the board
+(Figure~\ref{fig:closedloop}).
+
+\begin{figure}
+\centering
+
+\includegraphics[width=5.0in]{robot-closed-loop.eps}
+
+\caption{Block diagram of a high-level closed-loop autonomous system using \design{}.}
+\label{fig:closedloop}
+\end{figure}
 
@@
 
-% FIXME citation about why 2D is ``good enough''
 As only one-dimensional and two-dimensional population units have been implemented, networks that require populations to compute functions

### 2014-04-22 docs/paper/paper.tex
resource utilization

@@
 device (Xilinx XC7VX485T-2FFG1761C). This device was chosen because of its ready availability on a development platform and its large logic capacity and available resources.
-In particular, it provides 37080~Kb of on-chip block RAM and 2800 customizable DSP slices, both of which the Nengo-RT design requires in abundance in order to support
-the largest possible implementation size.
+In particular, it provides 37080~Kb of on-chip block RAM and 2800 customizable DSP slices, both of which the Nengo-RT design requires in abundance.
+Table~\ref{tbl:utilization} outlines the resource utilization of the implemented design.
 
-% FIXME implemented design size
+\begin{table}
+\begin{tabular}{|r|c|c|l|}
+\hline
+Resource & Used & Total & \% Used \\
+\hline
+\hline
+Registers & 82~149 & 607~200 & 13\% \\
+Lookup Tables & 101~859 & 303~600 & 33\% \\
+Logic Slices & 43~596 & 75~900 & 57\% \\
+Block RAM (18~Kb) & 914 & 2060 & 44\% \\
+DSP & 1722 & 2800 & 61\% \\
+\hline
+\end{tabular}
+\caption[\design{} resource utilization on VC707 development board.]
+{Resource utilization of the \design{} design with 7 one-dimensional and 7 two-dimensional population units, and one decoded value buffer reserved for external inputs.
+Resource totals are for the Xilinx XC7VX485T device.}
+\label{tbl:utilization}
+\end{table}
 

### 2014-04-22 docs/paper/paper.tex
power and temperature

@@
 both input and output are wait-free.
+
+Power analysis showed that the implemented design, with 7 one-dimensional and 7 two-dimensional population units,
+is estimated to use a maximum of 5.675~W of power while a simulation is running, assuming all population units are fully utilized.
+With the installed heatsink and fan on the VC707 board, the FPGA is expected to operate at an on-die temperature of $31.5\,^{\circ}{\rm C}$.
 
@@
 
-\subsection{power, \% timestep used for each simulation, \% of board resources for each simulation}
-
 \section{Discussion}

### 2014-04-22 docs/paper/paper.tex
real vs shared PCs

@@
 is estimated to use a maximum of 5.675~W of power while a simulation is running, assuming all population units are fully utilized.
+For an individual population unit, a one-dimensional unit consumes an average of 0.130~W of power during simulation,
+and a two-dimensional unit consumes an average of 0.425~W of power.
 With the installed heatsink and fan on the VC707 board, the FPGA is expected to operate at an on-die temperature of $31.5\,^{\circ}{\rm C}$.
@@
 populations with each set of principal components while minimizing the differences
-between the approximate principal components and the actual values. % FIXME explain/calculate why actual principal components cannot be used
+between the approximate principal components and the actual values.
+
+Populations must share principal components because storing a separate set of principal components for each population would require
+an unreasonably large amount of memory. As an example, a one-dimensional population currently uses seven principal components,
+each of which is a 1024-word, 12-bit block memory, for a total of 86~016~bits of memory per population.
+The FPGA on the VC707 development board has a maximum of approximately 37~080~000~bits of block memory;
+if we allowed each population to use its own principal components, we could not simulate more than 431 populations this way
+(and this number is halved for two-dimensional populations, which use 15 principal components).
+Therefore, in order to simulate large numbers of populations, it is necessary for large groups of populations to share principal components,
+even though this is not as accurate as providing a ``real'' set of principal components for each separate population.
+
 The compiler calculates decoders and encoders for each population with respect
@@
 
-\includegraphics[width=6in]{integrator-1d.eps} % FIXME software plot
+\includegraphics[width=6in]{integrator-1d.eps} % FIXME software plot (for all figures)
 % FIXME figure: larger axis and label font, make this consistent across all figures
@@
 oscillate in a sinusoidal fashion. This oscillation, once started, will proceed indefinitely.
-% FIXME frequency analysis?
 The oscillation as observed in the figure confirms the correct operation of the two-dimensional
@@
 \end{figure}
-
-% FIXME: explain input and expected output
 
@@
 which would require over 10000 times as many connections to provide all-to-all connectivity.
-% FIXME contrast this with size of neuron-to-neuron interconnect
-For large numbers
-of encoders and decoded value buffers, the complexity of the interconnect at
+For large numbers of encoders and decoded value buffers, the complexity of the interconnect at
 that design size makes it impossible for the FPGA design tools to find a way to

### 2014-06-21 docs/paper/paper.tex
software oscillator figure

@@
 
-\includegraphics[width=6in]{integrator-1d.eps} % FIXME software plot (for all figures)
+\includegraphics[width=6in]{integrator-1d.eps} % FIXME software plot
 % FIXME figure: larger axis and label font, make this consistent across all figures
@@
 
-\begin{subfigure}[b]{\textwidth}
+\begin{subfigure}[a]{\textwidth}
 \centering
@@
 
-\begin{figure}
+\begin{figure}[h!]
 \centering
-\includegraphics[width=6in]{oscillator-2d.eps}
 
-\caption[Simulation of a 2D neural oscillator.]
-{Hardware output from the simulation of a two-dimensional population representing a neural oscillator.}
+\begin{subfigure}[a]{0.4\textwidth}
+\centering
+\includegraphics[width=4in]{oscillator-2d.eps}
+\caption{Hardware simulation}
+\label{fig:oscillator2d:hw}
+\end{subfigure}
+
+\begin{subfigure}[b]{0.4\textwidth}
+\centering
+\includegraphics[width=4in]{oscillator-2d-sw.eps}
+\caption{Software simulation}
+\label{fig:oscillator2d:sw}
+\end{subfigure}
+
+\caption[Simulation of a 2D neural oscillator in hardware and software.]
+{Hardware (\ref{fig:oscillator2d:hw}) and software (\ref{fig:oscillator2d:sw}) output from the simulation of a two-dimensional population representing a neural oscillator.}
 \label{fig:oscillator2d}

### 2014-06-21 docs/paper/paper.tex
add integrator 1d figures

@@
 
-\includegraphics[width=6in]{integrator-1d.eps} % FIXME software plot
+\begin{subfigure}[a]{0.4\textwidth}
+\includegraphics[width=4in]{integrator-1d.eps}
+\caption{Hardware simulation}
+\label{fig:integrator1d:hw}
+\end{subfigure}
+
+\begin{subfigure}[b]{0.4\textwidth}
+\includegraphics[width=4in]{integrator-1d-sw.eps}
+\caption{Software simulation}
+\label{fig:integrator1d:sw}
+\end{subfigure}
+
 % FIXME figure: larger axis and label font, make this consistent across all figures
@@
 \caption[Simulation of a 1D neural integrator.]
-{Hardware output from simulation of a one-dimensional population representing a neural integrator.}
+{Hardware (\ref{fig:integrator1d:hw}) and software (\ref{fig:integrator1d:sw}) output from simulation of a one-dimensional population representing a neural integrator.}
 

### 2015-05-14 v1/v1.py
new 2D DOG code

@@ -1,0 +1,234 @@
+import numpy as np
+from scipy import ndimage as nd
+import matplotlib.pyplot as plt
+import pickle
+
+def make_DOG2(inner_sigma, x):
+    y = x
+    outer_sigma = inner_sigma*5
+    X, Y = np.meshgrid(x, y)
+    inner_gaussian = 1./(2.*np.pi*inner_sigma) * np.exp(-(X**2 + Y**2)/2./inner_sigma**2) 
+    outer_gaussian = 1./(2.*np.pi*outer_sigma) * np.exp(-(X**2 + Y**2)/2./outer_sigma**2) 
+#     return inner_gaussian - outer_gaussian
+    return inner_gaussian - outer_gaussian / 2
+
+def make_gabor(x, frequency, phase, sigma): 
+    """ Make a Gabor kernel. """
+    return np.cos(frequency*x + phase) * np.exp(-x**2/2./sigma**2)
+
+def make_gabors(x):
+    """
+    Creates 1D Gabor cross-sections 
+    
+    Returns:
+    -------
+    gaborx: 3xn cross-sections in sinusoid direction (different phases)
+    gabory: 1xn cross-section orthogonal to sinusoid direction    
+    """
+    gabor_frequency = 2*np.pi/30
+    gaborx = np.zeros((len(x), 3))
+    gaborx[:,0] = make_gabor(x, gabor_frequency, -np.pi/2, 10)
+    gaborx[:,1] = make_gabor(x, gabor_frequency, 0, 10)
+    gaborx[:,2] = make_gabor(x, gabor_frequency, np.pi/2, 10)
+    
+    y = x
+    gabory = np.exp(-y**2/2./8**2)
+    
+    return gaborx, gabory
+     
+def filter_dots(dots, DOG):
+    return nd.filters.convolve(dots, DOG)
+
+
+def shift_kernel(kernel, shape, centre):
+    h, w = kernel.shape
+    assert(h % 2 == 1)
+    assert(w % 2 == 1)
+    half_h = np.floor(h/2)
+    half_w = np.floor(w/2)
+    
+    result = np.zeros((shape[0]+2*half_h, shape[1]+2*half_w)) #zero pad to simplify edge handling 
+
+    ind_h = centre[0] + np.arange(0, 2*half_h+1, dtype='int')    
+    ind_w = centre[1] + np.arange(0, 2*half_w+1, dtype='int')
[…]

### 2015-05-14 docs/paper/paper.tex
fixed multiplier figures

@@
 \usepackage{listings}
+\usepackage{authblk}
+\usepackage{amsmath}
 
@@
 
+\title{Large-scale neural circuits on an FPGA}
+
+\author[1]{Murphy Berzish}
+\author[1,2,3]{Chris Eliasmith}
+\author[1,2]{Bryan Tripp}
+\affil[1]{Centre for Theoretical Neuroscience, University of Waterloo}
+\affil[2]{Dept. of Systems Design Engineering, University of Waterloo}
+\affil[3]{Dept. of Philosophy, University of Waterloo}
+\maketitle
+
+\abstract{abstract goes here - bryan???}
+
 \section{Introduction}
@@
 
-\subsection{Potential Applications}
+- potential applications
 
@@
 
+- from my other paper for mind collection
+
 \section{Methods}
 
-\subsection{Description of NEF and population mode}
+\subsection{The Neural Engineering Framework}
+
+The Neural Engineering Framework (NEF) is method for constructing biologically realistic neural models \cite{Eliasmith1999s}.
+Several overviews of the NEF are available \cite{Eliasmith2003f,Eliasmith2005p,stewart2012b,Eliasmith2012b}, but the description provided here is based on that in \cite{Stewart2014}.
+The NEF is a general-purpose approach to
+implementing algorithms using components such
+as spiking neurons \cite{Eliasmith2003m}. As a result, the NEF can be thought of
+as a ``neural compiler'' where algorithms characterized by a high-level
+mathematical description are converted into neural networks.
+This compilation process works for arbitrary neuron types, and can
+be constrained in biologically informed ways. Importantly, the high-level
+description must be expressed in terms of vectors and functions on
+those vectors (including differential equations). The resulting
+neural networks approximate the desired functions, and the error of
+this approximation can be made small by increasing the
+number of neurons.
+
+While the NEF can be used to build arbitrary abstract systems such
+as controlled attractor networks \cite{Eliasmith2005p}, we have primarily
+used it to show how particular capabilities found in real animals
+might be implemented biologically. This has included path integration
+in rodents \cite{Conklin2005b}, working memory \cite{Singh2006b}
+and arm movements \cite{Dewolf} in monkeys, and decision-making in
+rats \cite{Laubach2010,Liu2011} and humans \cite{Litt2008u}. We
+have also taken into account biological constraints such as Dale's
+Principle \cite{Parisien2008c} and incorporated biologically realistic
+learning rules to construct these networks \cite{MacNeil2011a,bekolay2013}.
+
+The remainder of this section serves as a summary of the three main principles
+of the NEF.
+
+\subsection{Principle 1 - Representation}
+
+In the NEF, groups of neurons are taken to represent vectors,
+and connections between groups of neurons compute functions on those
+vectors. The first NEF principle shows how the activity of a group
+of neurons can be said to represent a vector, and how changes in the
+activity of those neurons corresponds to changes in the represented
+vector.
+
+In the brain many neurons have a particular stimulus (or response) for
+which they will fire most strongly. As an input stimulus (or response)
+changes to become less similar to the preferred stimulus, the neuron
+will fire less quickly.  This kind of behaviour was originally identified in the motor
+system \cite{Georgopoulos1989q} and has since been seen in the head
+direction system \cite{Taube2007}, visual system \cite{Rust2006},
+and auditory system \cite{Fischer2009w}.
+The NEF identifies the preferred stimulus with a ``preferred direction vector'' associated with each neuron.
+For a more detailed exploration of this idea, see \cite{Stewart2011a}.
+
+In the NEF, we generalize this concept to all neural populations (those that are stimulus driven and those that are not).
+In particular, we quantify it by stating that the total current going
+into a neuron will be proportional to the dot product of the vector
+to be represented, $\mathbf{x}$, and the preferred direction (or ``encoding'') vector
+for the neuron, $\mathbf{e}_{i}$ (plus a constant bias term). The
+response of a neuron \textbf{$i$} for any given input vector \textbf{$\mathbf{x}$}
+is thus
+
+\begin{equation}
+\delta_{i}(\mathbf{x})=G_{i}[\alpha_{i}\mathbf{e}_{i}\mathbf{x}+J_{i}^{bias}]\label{eq:enc}
+\end{equation}
+where $\delta_{i}$ is the spiking output of the neuron, $G_{i}$
+is the neuron model, $\alpha_{i}$ is a randomly chosen gain term,
+$\mathbf{e}_{i}$ is the preferred direction vector, and $J_{i}^{bias}$
+is a randomly chosen fixed background current. We use $\mathbf{e}$
+for \textit{encoder} to indicate that (\ref{eq:enc}) captures a transformation
+between spaces: \textbf{$\mathbf{x}$} is encoded in the activity
+space of the neurons.  Importantly, the NEF applies
+to a large variety of neuron models (including both spiking and non-spiking
+models) since it makes no commitment to a specific function $G$ (whose
+input is the total current flowing into the neuron). The leaky-integrate-and-fire
+(LIF) model is a common choice, for reasons of computational efficiency, but a wide variety of
+neural models work with the NEF.
+
+Given an encoding operation, it is natural to define a \textit{decoding} operation,
+in order to characterize the information processing characteristics
+of the system (in this case an ensemble of neurons). To decode a continuous
+estimate of the input from neural activity, the NEF focuses on the
+postsynpatic filtered activity generated by the reception of a spike
+at a synapse. This activity is taken to be a linear filter applied
+to the spiking activity:
+\[
+a_{i}(\mathbf{x})=\sum_{j}h_{i}(t)*\delta_{i}(t-t_{j}(\mathbf{x}))
+\]
+where $h_{i}(t)$ is the synaptic response function (usually a decaying
+exponential) with a time constant, $\tau_{PSC}$, determined by neurotransmitter
+type at the synapse, '$*$' is the convolution operator, and $\delta_{i}(t-t_{j}(\mathbf{x}))$
+is the spike train produced by neuron $i$ in response to input $\mathbf{x}$,
+with spike times indexed by $j$. Having defined this continuous variable,
+which is equal to the spike rate as $\tau_{PSC}\rightarrow\infty$,
+we can specify a decoding operation for estimating the input $\mathbf{x}$.
+
+For reasons that will be apparent in a moment, we use a \textit{linear}
+decoder:
+
+\begin{equation}
+\hat{\mathbf{x}}=\sum_{i}^{N}a_{i}\mathbf{(x)}\mathbf{d}_{i}\label{eq:dec}
+\end{equation}
+where $N$ is the number of neurons in the group, $\mathbf{d}_{i}$
+are the linear decoders, and $\hat{\mathbf{x}}$ is the estimate of
+the original $\mathbf{x}$ value that produced the neural activity
+(\ref{eq:enc}).
+
+Any optimization method can be used to find these decoders. The simplest
+is to use standard least-squares optimization:
+\begin{equation}
+\underset{\mathbf{d}_{i}}{arg\, min}\int[\mathbf{x}-\sum_{i}^{N}a_{i}(\mathbf{x)}\mathbf{d}_{i}]^{2}d\mathbf{x}\label{eq:error}
+\end{equation}
+where $\mathbf{d}_{i}$ are the decoding vectors over which this error
+is minimized and the integral is over all $\mathbf{x}$ values. It
+has been shown that linear decoders are sufficient to decode \textasciitilde{}95\%
+of the information available in a single spike train generated by
+a stimulus \cite{Rieke1997b}. Furthermore, as the number of neurons
+$N$ increases, the mean squared error decreases as $1/N$. The NEF
+decoding process is depicted in Figure \ref{fig:NEF-decoding}, where
+the optimal linear decoders have been found and used for twenty neurons.
+Temporal decoding is also linear, as described, and is performed using
+the postsynaptic, current-based filters (see section \ref{sub:Principle-3--}
+for further discussion).
+
+While linear decoders are useful for visualizing the information encoded
+within the activity of a group of neurons, they also provide a direct
+way to solve for connection weights between groups of neurons. This
+is a key advantage of the NEF: rather than using a learning rule to
+optimize over the entire space of all connection weights, we instead
+solve the simpler problem of optimizing over the space of linear decoders,
+and then use that result to solve for the connection weights. Importantly,
+given the characterization of neural activity with preferred direction
+vectors, there is no difference between using this smaller space and
+using the equivalent full connection matrix. Interestingly, this general
+technique has started to be applied in broader domains; see \cite{tapson2013}
+for an overview.
+
+For example, if a connection between neural groups is meant to compute
+the identity function $\mathbf{y}=\mathbf{x}$ (where $\mathbf{y}$
+is the vector space represented by the second population B and $\mathbf{x}$
+is the vector space represented by the first population A), the connections
+between individual neurons are given by 
+\begin{equation}
+\omega_{ij}=\alpha_{j}\mathbf{d}_{i}\mathbf{e}_{j}\label{eq:weights}
+\end{equation}
+where $i$ indexes the neurons in group A and $j$ indexes the neurons
+in B. Behavior of this network is shown in Figure \ref{fig:Using-NEF-weights}a.
+
+While the least-squares method for optimization we use here is not
+biologically plausible on its own, we have also shown that biologically
+realistic learning rules will converge on a similar solution \cite{MacNeil2011a}.
+These realistic rules are, however, several orders of magnitude more
+computationally expensive.
+
+\subsection{Principle 2 - Transformation}
+
+Connections between groups of neurons can also compute functions other
+than the identity function. That is instead of \textbf{$\mathbf{y}=\mathbf{x}$}
+we can do \textbf{$\mathbf{y}=f(\mathbf{x})$}. We do this by finding
+decoders \textbf{$\mathbf{d}_{i}^{f}$ }for the particular function\textbf{
+$f(\mathbf{x})$} by substituting \textbf{$f(\mathbf{x})$ }for $\mathbf{x}$
+in equation (\ref{eq:error}).
+
+\begin{equation}
+\underset{\mathbf{d}_{i}^{f}}{arg\, min}\int[f(\mathbf{x})-\sum_{i}a_{i}(\mathbf{x)}\mathbf{d}_{i}^{f}]^{2}d\mathbf{x}.\label{eq:error-fcns}
+\end{equation}
+
+The connection weights can then be computed using (\ref{eq:weights}).
+For the special case of linear functions\textbf{ $\mathbf{y}=\mathbf{L}\mathbf{x}$},
+rather than solving for a new decoder, we can simply put $\mathbf{L}$
+directly into the weight equation itself, resulting in $\omega_{ij}=\alpha_{j}\mathbf{d}_{i}\mathbf{L}\mathbf{e}_{j}$.
+Combining these two approaches, the neural connection weights needed
+to approximate the function\textbf{ $\mathbf{y=L}f(\mathbf{x\textrm{)}}$}
+are:
+\begin{equation}
+\omega_{ij}=\alpha_{j}\mathbf{d}_{i}^{f}\mathbf{L}\mathbf{e}_{j}\label{eq:weights-general}
+\end{equation}
+Figure \ref{fig:Using-NEF-weights}b shows the computation of the
+element-wise square ($f(\mathbf{x})=[x_{1}^{2},x_{2}^{2}]$).
+
+\subsection{\label{sub:Principle-3--}Principle 3 - Dynamics}
+
+While the first two principles are sufficient to build neural approximations
+of any desired function of the vector $\mathbf{x}$, the NEF also
+provides a method for computing functions of the form \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}f(\mathbf{x},\mathbf{u})$},
+where $\mathbf{u}$ is the input from some other population. We do
+this by exploiting the fact that neurons do not simply accept input
+as spikes. Rather, when a spike is transmitted from one neuron to
+another, the actual current that flows into the second neuron is a
+low-pass-filtered version of that spike. In particular, the post-synaptic
+current is well-approximated by $h(t)=u(t)e^{-t/\tau}$, where $u(t)$
+is the step function and $\tau$ is the time constant of the neurotransmitter
+used. This time constant varies throughout the brain, e.g., from 2-5
+ms (AMPA; \cite{jonas1993quantal}) up to \textasciitilde{}100ms (NMDA;
+\cite{sah1990properties}). The effect of this filter is that instead
+of a connection computing the function $\mathbf{y}(t)=f(\mathbf{x}(t))$
+it will compute $\mathbf{y}(t)=f(\mathbf{x}(t))*h(t)$ (or, in the
+Laplace domain, $\mathbf{Y}(s)=\mathbf{F}(s)H(s)$).
+
+It turns out that this implicit low-pass filter can be used to generate
+neural models that compute abitrary dynamics. Given a neural population
+representing $\mathbf{x}$, an input $\mathbf{u}(t)$, and a feedback
+connection from $\mathbf{x}$ back to itself computing $g(\mathbf{x}(t))$,
+we note that in the Laplace domain we get $\mathbf{X}(s)=(\mathbf{G}(s)+\mathbf{U}(s))H(s)$.
+Since the Laplace transform of $h(t)=u(t)e^{-t/\tau}$ is $H(s)=1/(1+s\tau)$,
+we can rearrange this to get $s\mathbf{X}(s)=(\mathbf{G}(s)-\mathbf{X}(s))/\tau+\mathbf{U}(s)/\tau$.
+Converting back to the time domain, \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}\frac{g(\mathbf{x}(t))-\mathbf{x}(t)}{\tau}+\frac{\mathbf{u}(t)}{\tau}$}.
+Thus, if we desire the dynamics \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}f(\mathbf{x}(t))+\mathbf{u}(t)$},
+we introduce a feedback connection that uses the previous two NEF
+principles to find connection weights that compute $g(\mathbf{x}(t))=\tau f(\mathbf{x})+\mathbf{x}$
+and we scale the input $\mathbf{u}(t)$ by $\tau$. For arbitrary
+dynamics of the form \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}f(\mathbf{x},\mathbf{u})$},
+we do a change of variables $\mathbf{x}^{\prime}=\langle\mathbf{x},\mathbf{u}\rangle$.
+
+This exploitation of the inherent first-order low-pass filter found
+in synaptic connections allows for the implementation of a very wide
+variety of systems, including oscillators, integrators, and arbitrary
+attractor networks \cite{Eliasmith2005p}. Figure \ref{fig:Two-dynamical-systems}a
+shows a single neural population with a recurrent feedback connection
+computing the standard linear oscillator \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}[-x_{2},-x_{1}]$}.
+Here, the post-synapic time constant $\tau$ is 100ms, but the period
+of oscillation is $2\pi$ (\textasciitilde{}6.28 seconds). Importantly,
+we achieve different periods of oscillation without changing the neural
+property $\tau$; rather, we change the function being computed by
+the feedback connection, resulting in a different set of connection
+weights. For example, we could compute \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}[-x_{2}/2,-x_{1}/2]$},
+giving a period of $\pi$, instead. This same method works for nonlinear
+functions as well; in Figure \ref{fig:Two-dynamical-systems}b we
+show the classic Lorenz attractor \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}[10(x_{2}-x_{1}),x_{1}(28-x_{3})-x_{2},x_{1}x_{2}-\frac{8}{3}x_{3}]$}.
+This approach allows for the construction of neural models that correspond
+to a very large family of functions, including those typically employed
+by modern control theory and dynamic systems theory.
+
+It should be noted that networks constructed with this approach are
+fundamentally different from Echo State Networks \cite{jaeger2001}
+and Liquid State Machines \cite{Maass2002}. In both of those approaches,
+the recurrent connections are randomly chosen, and then a linear combination
+of the outputs are found that compute the desired function. With the
+NEF, we also find a linear combination of the outputs ($\mathbf{d}_{i}$),
+but we solve for the ideal recurrent connection weights to achieve
+the desired dynamics. This makes the NEF much more efficient and capable
+of computing a much wider range of dynamics.
+
+\subsection{Nengo}
+
+The principles of the NEF have been incorporated into the Nengo neural simulator \cite{Bekolay2014}.  This simulator
+
+- mention spaun
+
+These three principles are sufficient to implement all of our neural
+models. However, to simplify the process of constructing these models,
+we have developed an open-source software package known as Nengo (Neural
+ENGineering Objects) that creates and runs these models. Models can
+be built in Nengo using a drag-and-drop graphical user interface or
+specified using the Python scripting language. Full details and documentation
+can be found online at http://nengo.ca, and in other publications
+\cite{Stewart2009l}.
+
+For example, to create the model shown in Figure \ref{fig:Using-NEF-weights}a,
+we use the following script:
+\begin{verbatim}
+net = nef.Network("Identity Function")
+net.make("A", neurons=20, dimensions=1) 
+net.make("B", neurons=20, dimensions=1)
+net.connect("A", "B") 
+\end{verbatim}
+
+For Figure \ref{fig:Using-NEF-weights}b we need to compute the element-wise
+square. This is specified in a Python function as follows:
+
+\begin{verbatim}
+net = nef.Network("Element-wise Square")
+net.make("A", neurons=20, dimensions=1) 
+net.make("B", neurons=20, dimensions=1)
+
+def square(x):
+    return x{[}0{]}{*}x{[}0{]}, x{[}1{]}{*}x{[}1{]}
+
+net.connect("A", "B", func=square) 
+\end{verbatim}
+
+Nengo will automatically solve for the connection weights that will
+best approximate the provided function.
+
+Nengo also provides an interactive interface for displaying the results
+of a simulation while it is running, allowing for real-time interaction
+with a running model. This interface allows the generated plots to
+be exported, and was used to produce the previous figures in this
+paper (i.e. all figures in section \ref{sec:The-Neural-Engineering}).
+Furthermore, Nengo scales up to our largest models: the 2.5 million
+neuron Spaun model is run in Nengo, and can be downloaded at http://models.nengo.ca/spaun.
+
+- backends
+
+\subsubsection{Population mode}
+
+- Bryan???
 
 \subsubsection{Considerations related to hardware (multiplexing, clustering, using principal components)}
+
+- Murphy???
 
@@
 
-\begin{figure}
+\begin{figure}[h!]
 \centering
@@
 
-\begin{subfigure}[b]{0.5\textwidth}
+\begin{subfigure}[b]{0.4\textwidth}
 \centering
-\includegraphics[width=3in]{multiplier-sw.eps}
+\includegraphics[width=4in]{multiplier-sw.eps}
 \caption{Software}
@@
 
-\begin{subfigure}[b]{0.5\textwidth}
+\begin{subfigure}[b]{0.4\textwidth}
 \centering
-\includegraphics[width=3in]{multiplier.eps}
+\includegraphics[width=4in]{multiplier.eps}
 \caption{Hardware}
@@
 
+\bibliographystyle{acm}
+\bibliography{references}
+
 \end{document}

### 2015-05-14 v1/v1.py
adding 10x10 array of receptive fields

@@ -219,13 +219,24 @@ def get_default_inputs():
 #     filtered = pickle.load(open("/Users/bptripp/code/nengo-FPGA/v1/dog-filtered-dots.p", "rb" ))
     w_kernel = pickle.load(open("/Users/bptripp/code/nengo-FPGA/v1/dog-gabor-weights-new-2.p", "rb" ))
     filtered = pickle.load(open("/Users/bptripp/code/nengo-FPGA/v1/dog-filtered-dots-new.p", "rb" ))
-    centres = np.array([[200,200]]) 
+    
+    c = np.arange(87, 313, 25)
+    centres = np.zeros((c.size**2,2), dtype='int')
+    for i in range(c.size): 
+        for j in range(c.size): 
+            centres[i*c.size+j,0] = c[i]
+            centres[i*c.size+j,1] = c[j]
+    
+    print(centres)
+    
+#     centres = np.array([[200,200]]) 
     return get_inputs(w_kernel, filtered, centres)
 
 
 # get_default_weights()
 
 inputs = get_default_inputs()
+pickle.dump(inputs, open("/Users/bptripp/code/nengo-FPGA/v1/inputs.p", "wb" ))
 # inputs = pure_gabor()
 plt.plot(inputs[0,0,:]-10, 'r')
 plt.plot(inputs[0,1,:], 'g')

### 2015-05-14 docs/paper/paper.tex
added first two sections
-finished NEF description
- finished intro
- finished neuromorphic section

@@
 \usepackage{amsmath}
+\usepackage{url}
+\usepackage{marvosym}
 
-\newcommand{\design}{Nengo-RT}
+\newcommand{\design}{Nengo\char`_FPGA}  %% Could be xFPGA to specify xilinx? since there might be other fpga backends
 
@@
 
+% I don't care about order, just assumed you'd want to be 'senior author'
 \author[1]{Murphy Berzish}
@@
 
-\abstract{abstract goes here - bryan???}
+\begin{abstract}
+abstract goes here - bryan???
+\end{abstract}
 
@@
 
-\subsection{Motivation}
+Large-scale neural models have recently become centrepieces of very large international research investments.
+The controversial European Human Brain Project (HBP) has a central goal of developing a human-scale neural simulation in the next 10 years, and has been granted \EURtm 1 billion to do so.
+Similarly, the Obama Brain Initiative (OBI) in the US is investing well over \$1 billion over a similar timeframe in theoretical and experimental neuroscience.
+Both projects have identified massive increases in computational performance as a critical for achieving their goals.
 
-- potential applications
+One approach to solving this challenge is employing large supercomputers, such as the IBM BlueGene.
+Both HBP and OBI have exploited this resource to simulation millions or billions of neurons, although not in real time.
+However, such resources are available to only a select few researchers.
 
-\subsection{State of the Art (Neuromorphic Hardware)}
+Another approach is to build specialized hardware -- often called ``neuromorphic'' hardware -- to massively increase the efficiency and speed of neural simulations.
+There have been several exciting developments in this regards, both in industry and in academia.
+IBM recently announced what will soon be the first commercially available neuromorphic chip \cite{Merolla2014}.
+As well, several academic projects, such as the SpiNNaker \cite{Khan2008} and Neurogrid \cite{Choudhary2012} projects have made impressive progress in recent years (see Section \ref{sub:neuromorphic} for further discussion).
+However, these solutions are not available in large quantities, and hence remain inaccessible to a majority of researchers.
 
-- from my other paper for mind collection
+Here, we focus on developing an affordable and accessible solution to the problem of simulating large-scale neural circuits.
+Our approach allows anyone with access to off-the-shelf FPGA hardware to simulate models up to about 10 million equivalent neurons with arbitrary connectivity in real time.
+
+Importantly, this level of scaling is dependent on the notion of ``equivalent neurons.'' In Section {\ref{sub:NEF}} we discuss our theoretical approach that allows extremely good approximations of the activity of about a thousand spiking neurons by simulating them as a single population, at a resolution of about 1~ms.
+Simulating large neural circuits at this level of abstraction allows for very computationally efficient solutions that preserve function.
+The main cost of this excellent computational scaling is that these methods do not allow online tuning of the connection weights between individual cells.
+However, a large-class of adaptive and non-adaptive massively parallel circuits can be effectively simulated, providing a useful tool for exploring the class of brain-like algorithms.
+
+Indeed, one of the main benefits of our approach is that it focusses both on function and biological plausibility, in contrast to work presented by HBP and OBI  which captures little function \cite{eliasmith2013a}.
+This focus remains here, allowing the methods presented to be applied both to characterizing biological cognition and to exploring engineering applications of brain-like algorithms.
+To this end, we present examples that compute specific mathematical functions, as well as those that simulate early visual processing of time-varying stimuli.
+
+We begin our discussion by describing related neuromorphic approaches in the next section.
+We then explain our methods for constructing neural simulations, and demonstrate the equivalence between single cell and population-level simulations.
+Next, we discuss the hardware design and show that it is able to efficiently simulate a variety of small test circuits.
+We then show that the system scales efficiently to simulating very large models in real time.
+We conclude by discussing possible improvements to the current design.
+
+\subsection{\label{sub:neuromorphic}Current Neuromorphic Hardware\protect\footnote{This discussion is based on that in \cite{eliasmith2015a}.}}
+
+Perhaps best known neuromorphic hardware is that being developed by IBM.
+The architecture itself is called TrueNorth.
+It includes a digital model of individual neurons, and a method for programming the architecture \cite{Esser2013}.
+The purpose of this architecture is to exploit brain-like organization, to permit the simulation of a specific class of algorithms that do not efficiently run on classical computer architectures.
+IBM researchers argue that the ``low-precision, synthetic, simultaneous, pattern-based metaphor of TrueNorth is a fitting complement to the high-precision, analytical, sequential, logic-based metaphor of today's of von Neumann computers'' \cite[p.~1]{Esser2013}.
+
+TrueNorth has neurons organized into 256 neuron blocks in which each neuron can receive input from 256 axons.
+To assist with programming this hardware, IBM has introduced the notion of a ``corelet'', which is an abstraction that encapsulates local connectivity in small networks.
+These act like small programs that can be composed in order to build up more complex functions.
+To date the demonstrations of the approach have focused on simple, largely feed-forward, standard applications, although they have employed wide range of methods including RBMs, liquid state machines, HMMs, and very simple NEF networks.
+It should be noted that the proposed chip does not yet exist, and current demonstrations are on detailed simulations of the architecture.
+However, because it is a digital chip the simulations are highly accurate.
+
+A direct competitor to IBM's approach is the Zeroth neuromorphic chip from Qualcomm.
+Like IBM, Qualcomm believes that constructing brain-inspired hardware will provide a new paradigm for exploiting the efficiencies of neural computation for the kinds of information processing that brains excel at, but which are extremely challenging for standard von Neumann approaches.
+The main difference between these 2 approaches is that Qualcomm has committed to allowing online learning to take place on the hardware.
+Consequently, they announced their processor by demonstrating its application in a reinforcement learning paradigm on a real-world robot.
+They have released videos of the robot maneuvering in an environment and learning to only visit one kind of stimulus (white boxes; \url{http://www.youtube.com/watch?v=8c1Noq2K96c}).
+It should again be noted that this is an FPGA simulation of a digital chip which does not yet exist.
+However the simulation, like IBM's, is highly accurate.
+However, it has recently been suggested that Qualcomm is no longer fully supporting this research effort.
+
+Both of these projects are spearheaded by industry.
+In the academic sphere, the Spinnaker project at Manchester University has not focused on designing new kinds of chips, but instead focused on using low-power ARM processors at a massive scale to allow large-scale brain simulations \cite{Khan2008}.
+As a result, the focus has been on designing approaches to routing that allow for the high bandwidth communication that underwrites much of the brain's information processing.
+Simulations on the Spinnaker hardware typically employ spiking neurons, like IBM and Qualcomm, and occasionally allow for learning \cite{davies2013spike}, like Qualcomm's approach.
+However, even with low power conventional chips, the energy usage is projected to be higher on the Spinnaker platform per neuron.
+Nevertheless, Spinnaker boards have been used in a wider variety of larger-scale embodied, and non-embodied applications.
+These include simulating place cells, path integration, item classification, and simple sensory guided movements \cite{Stewart2013a}.
+
+There are also a number of neuromorphic projects that use analog instead of digital implementations of neurons.
+Analog approaches tend to be several orders of magnitude more power efficient \cite{hasler2013}, although also more noisy, unreliable, and subject to process variation.
+These projects include work on the Neurogrid chip at Stanford University \cite{Choudhary2012}, and on a chip at ETH Z\"{u}rich \cite{Corradi2014}.
+The Neurogrid chip has demonstrated large numbers of simulated neurons, up to 1 million, while the ETH Z\"{u}rich chip allows for online learning.
+More recently, the Neurogrid chip has been used to control a nonlinear, 6 degree of freedom robotic arm, exhibiting perhaps the most sophisticated information processing from an analog chip to date \cite{Menon14:biorob}.
+
+In addition to the above neuromorphic projects, which are focused on cortical simulation, there have been several specialized neuromorphic chips that mimic the information processing of different perceptual systems.
+For example, the dynamic vision sensor (DVS) artificial retina developed at ETH Z\"{u}rich performs real-time vision processing that results in a stream of neuron-like spikes \cite{Lichtsteiner2008}.
+Similarly, an artificial cochlea called AEREAR2 has been developed that generate spikes in response to auditory signals \cite{Li2012}.
+The development of these and other neuronal sensors make it possible to build fully embodied spiking neuromorphic systems \cite{Galluppi2014}.
+
+More directly relevant to the present paper, there have be several attempts to build neuromorphic systems on FPGAs.
+For instance, \cite{cassidy2011design} describe a design with a million ``neurons'', where each neuron is abstracted as an arithmetic logic unit.
+In contrast to our design, all neurons are considered identical, there are no synaptic dynamics, and there is no modelling of neural noise.
+In \cite{li2012a}, synaptic dynamics are included, but the network consists of a total of 256 neurons.
+More recently, in \cite{wang2013} a network of over 4 thousand neurons is implemented.
+However, the approach we describe here is focussed on simulations of much larger numbers of effective neurons.
 
@@
 
-\subsection{The Neural Engineering Framework}
+\subsection{\label{sub:NEF}The Neural Engineering Framework}
 
@@
 in order to characterize the information processing characteristics
-of the system (in this case an ensemble of neurons). To decode a continuous
+of the system. To decode a continuous
 estimate of the input from neural activity, the NEF focuses on the
-postsynpatic filtered activity generated by the reception of a spike
+postsynaptic filtered activity generated by the reception of a spike
 at a synapse. This activity is taken to be a linear filter applied
@@
 which is equal to the spike rate as $\tau_{PSC}\rightarrow\infty$,
-we can specify a decoding operation for estimating the input $\mathbf{x}$.
-
-For reasons that will be apparent in a moment, we use a \textit{linear}
-decoder:
+we can specify a decoding operation for estimating the input $\mathbf{x}$:
 
@@
 
-Any optimization method can be used to find these decoders. The simplest
-is to use standard least-squares optimization:
+Any optimization method can be used to find these decoders. A simple example is least-squares optimization:
 \begin{equation}
@@
 where $\mathbf{d}_{i}$ are the decoding vectors over which this error
-is minimized and the integral is over all $\mathbf{x}$ values. It
-has been shown that linear decoders are sufficient to decode \textasciitilde{}95\%
-of the information available in a single spike train generated by
-a stimulus \cite{Rieke1997b}. Furthermore, as the number of neurons
-$N$ increases, the mean squared error decreases as $1/N$. The NEF
-decoding process is depicted in Figure \ref{fig:NEF-decoding}, where
-the optimal linear decoders have been found and used for twenty neurons.
-Temporal decoding is also linear, as described, and is performed using
-the postsynaptic, current-based filters (see section \ref{sub:Principle-3--}
-for further discussion).
+is minimized and the integral is over all $\mathbf{x}$ values. As the number of neurons
+$N$ increases, the mean squared error decreases as $1/N$.
 
-While linear decoders are useful for visualizing the information encoded
-within the activity of a group of neurons, they also provide a direct
-way to solve for connection weights between groups of neurons. This
+Note that employing linear decoding allows us to compute connection weights. This
 is a key advantage of the NEF: rather than using a learning rule to
@@
 solve the simpler problem of optimizing over the space of linear decoders,
-and then use that result to solve for the connection weights. Importantly,
-given the characterization of neural activity with preferred direction
-vectors, there is no difference between using this smaller space and
-using the equivalent full connection matrix. Interestingly, this general
-technique has started to be applied in broader domains; see \cite{tapson2013}
-for an overview.
-
-For example, if a connection between neural groups is meant to compute
+and then use that result to solve for the connection weights. For example, if a connection between neural groups is meant to compute
 the identity function $\mathbf{y}=\mathbf{x}$ (where $\mathbf{y}$
@@
 where $i$ indexes the neurons in group A and $j$ indexes the neurons
-in B. Behavior of this network is shown in Figure \ref{fig:Using-NEF-weights}a.
+in B.
 
-While the least-squares method for optimization we use here is not
-biologically plausible on its own, we have also shown that biologically
-realistic learning rules will converge on a similar solution \cite{MacNeil2011a}.
+While the least-squares method for optimization is not
+biologically plausible on its own, we have shown that biologically
+realistic learning rules will converge to a similar solution \cite{MacNeil2011a}.
 These realistic rules are, however, several orders of magnitude more
@@
 
-Connections between groups of neurons can also compute functions other
-than the identity function. That is instead of \textbf{$\mathbf{y}=\mathbf{x}$}
-we can do \textbf{$\mathbf{y}=f(\mathbf{x})$}. We do this by finding
+Principle 1 effectively assumes that we want to compute the identity function. Connections between groups of neurons can also compute functions other
+than the identity function. That is, instead of \textbf{$\mathbf{y}=\mathbf{x}$}
+we can compute \textbf{$\mathbf{y}=f(\mathbf{x})$}. We do this by finding
 decoders \textbf{$\mathbf{d}_{i}^{f}$ }for the particular function\textbf{
@@
 \end{equation}
-Figure \ref{fig:Using-NEF-weights}b shows the computation of the
-element-wise square ($f(\mathbf{x})=[x_{1}^{2},x_{2}^{2}]$).
 
@@
 While the first two principles are sufficient to build neural approximations
-of any desired function of the vector $\mathbf{x}$, the NEF also
+of any desired feedforward function of the vector $\mathbf{x}$, the NEF also
 provides a method for computing functions of the form \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}f(\mathbf{x},\mathbf{u})$},
-where $\mathbf{u}$ is the input from some other population. We do
-this by exploiting the fact that neurons do not simply accept input
+where $\mathbf{u}$ is the input from some other population. The NEF exploits the fact that neurons do not simply accept input
 as spikes. Rather, when a spike is transmitted from one neuron to
@@
 used. This time constant varies throughout the brain, e.g., from 2-5
-ms (AMPA; \cite{jonas1993quantal}) up to \textasciitilde{}100ms (NMDA;
+ms (AMPA; \cite{jonas1993quantal}) up to $\sim$100ms (NMDA;
 \cite{sah1990properties}). The effect of this filter is that instead
@@
 It turns out that this implicit low-pass filter can be used to generate
-neural models that compute abitrary dynamics. Given a neural population
+neural models that compute arbitrary dynamics. Given a neural population
 representing $\mathbf{x}$, an input $\mathbf{u}(t)$, and a feedback
@@
 variety of systems, including oscillators, integrators, and arbitrary
-attractor networks \cite{Eliasmith2005p}. Figure \ref{fig:Two-dynamical-systems}a
-shows a single neural population with a recurrent feedback connection
-computing the standard linear oscillator \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}[-x_{2},-x_{1}]$}.
-Here, the post-synapic time constant $\tau$ is 100ms, but the period
-of oscillation is $2\pi$ (\textasciitilde{}6.28 seconds). Importantly,
-we achieve different periods of oscillation without changing the neural
-property $\tau$; rather, we change the function being computed by
-the feedback connection, resulting in a different set of connection
-weights. For example, we could compute \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}[-x_{2}/2,-x_{1}/2]$},
-giving a period of $\pi$, instead. This same method works for nonlinear
-functions as well; in Figure \ref{fig:Two-dynamical-systems}b we
-show the classic Lorenz attractor \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}[10(x_{2}-x_{1}),x_{1}(28-x_{3})-x_{2},x_{1}x_{2}-\frac{8}{3}x_{3}]$}.
+attractor networks \cite{Eliasmith2005p}. 
 This approach allows for the construction of neural models that correspond
@@
 
-It should be noted that networks constructed with this approach are
-fundamentally different from Echo State Networks \cite{jaeger2001}
-and Liquid State Machines \cite{Maass2002}. In both of those approaches,
-the recurrent connections are randomly chosen, and then a linear combination
-of the outputs are found that compute the desired function. With the
-NEF, we also find a linear combination of the outputs ($\mathbf{d}_{i}$),
-but we solve for the ideal recurrent connection weights to achieve
-the desired dynamics. This makes the NEF much more efficient and capable
-of computing a much wider range of dynamics.
-
 \subsection{Nengo}
 
-The principles of the NEF have been incorporated into the Nengo neural simulator \cite{Bekolay2014}.  This simulator
+The principles of the NEF have been incorporated into the Nengo neural simulator \cite{Bekolay2014}.
+Nengo is a general purpose simulator that incorporates the three principles of the NEF.
+It has been used to build what is currently the world's largest functional brain model \cite{Eliasmith2012b}, along with many other simulations.
+Full details and documentation can be found online at \url{http://github.com/nengo/nengo}.
 
-- mention spaun
+One central feature of Nengo is that it permits the same model description to target a variety of ``backends.''  Current backends include CPUs, GPUs, the IBM BlueGene supercomputer, and several neuromorphic chips, including SpiNNaker \cite{Khan2008} and Neurogrid \cite{Choudhary2012}.
+Because Nengo exploits the NEF methods, model descriptions can be very concise.
+Model descriptions are written in python, although backends compile to a variety of languages, depending on the target hardware.
+For example, to create a recurrent model of a chaotic dynamical system (a Lorenz attractor) with 2000 spiking neurons and run it for 10 seconds, the following suffices:
 
-These three principles are sufficient to implement all of our neural
-models. However, to simplify the process of constructing these models,
-we have developed an open-source software package known as Nengo (Neural
-ENGineering Objects) that creates and runs these models. Models can
-be built in Nengo using a drag-and-drop graphical user interface or
-specified using the Python scripting language. Full details and documentation
-can be found online at http://nengo.ca, and in other publications
-\cite{Stewart2009l}.
+\begin{verbatim}
+import nengo
 
-For example, to create the model shown in Figure \ref{fig:Using-NEF-weights}a,
-we use the following script:
-\begin{verbatim}
-net = nef.Network("Identity Function")
-net.make("A", neurons=20, dimensions=1) 
-net.make("B", neurons=20, dimensions=1)
-net.connect("A", "B") 
+model = nengo.Network()
+with model:
+    state = nengo.Ensemble(2000, 3, radius=60)
+
+	def feedback(x):
+	    dx0 = -10 * x[0] + 10 * x[1]
+	    dx1 = -x[0] * x[2] - x[1]
+	    dx2 = x[0] * x[1] - 8.0/3 * (x[2] + 28) - 28
+
+	    return [dx0 * 0.1 + x[0],
+	            dx1 * 0.1 + x[1],
+	            dx2 * 0.1 + x[2]]
+
+    nengo.Connection(state, state, function=feedback, synapse=0.1)
+    state_probe = nengo.Probe(state, synapse=tau)
+
+sim = nengo.Simulator(model)
+sim.run(10)
 \end{verbatim}
 
-For Figure \ref{fig:Using-NEF-weights}b we need to compute the element-wise
-square. This is specified in a Python function as follows:
+Nengo will use the NEF methods to automatically solve for the connection weights that will
+best approximate the provided nonlinear, dynamic function.  To run this same network on a different backend, we only need change the second last line to indicate which simulator to use.
 
-\begin{verbatim}
-net = nef.Network("Element-wise Square")
-net.make("A", neurons=20, dimensions=1) 
-net.make("B", neurons=20, dimensions=1)
+Nengo also provides an interactive interface for building models, and displaying the results of a simulation while it is running, allowing for real-time interaction
+with a running model.
 
-def square(x):
-    return x{[}0{]}{*}x{[}0{]}, x{[}1{]}{*}x{[}1{]}
-
-net.connect("A", "B", func=square) 
-\end{verbatim}
-
-Nengo will automatically solve for the connection weights that will
-best approximate the provided function.
-
-Nengo also provides an interactive interface for displaying the results
-of a simulation while it is running, allowing for real-time interaction
-with a running model. This interface allows the generated plots to
-be exported, and was used to produce the previous figures in this
-paper (i.e. all figures in section \ref{sec:The-Neural-Engineering}).
-Furthermore, Nengo scales up to our largest models: the 2.5 million
-neuron Spaun model is run in Nengo, and can be downloaded at http://models.nengo.ca/spaun.
-
-- backends
+In this paper, we describe an extension to the Nengo simulator that allows models to compile to an FPGA, using a simulation method that efficiently approximates the computations performed by populations of spiking neurons.
 
@@
 - Bryan???
+- mention the other modes first to provide context/more info about Nengo
 
@@
 
-\design{} is a custom digital hardware design that implements the population-mode NEF in real-time (1~ms timestep).
+\design{} is a custom digital hardware design that implements the population-mode NEF in real-time (1~ms timestep) on an FPGA.
 A block diagram of the design is shown in Figure~\ref{fig:system}.
@@
 Nengo supports a more advanced connection called a learning connection,
-which mdoifies the decoders of a population according to a particular ``learning rule''
+which modifies the decoders of a population according to a particular ``learning rule''
 as a function of other represented values. The hardware design does not support this special connection type.

### 2015-05-15 docs/paper/paper.tex
adding population mode section to paper

@@
 abstract goes here - bryan???
+- ??? I think we want spiking neuron and population model simulation comparisons for the software simulations when comparing to hardware
 \end{abstract}
@@
 
+% @Chris: I don't get the last sentence above -- wasn't everything fine with x and u?
+
 This exploitation of the inherent first-order low-pass filter found
@@
 
-In this paper, we describe an extension to the Nengo simulator that allows models to compile to an FPGA, using a simulation method that efficiently approximates the computations performed by populations of spiking neurons.
+In this paper, we describe an extension to the Nengo simulator that allows models to compile to an FPGA, using a simulation method that efficiently approximates the computations performed by populations of spiking neurons (see "population mode", next section).
 
@@
 
-- Bryan???
-- mention the other modes first to provide context/more info about Nengo
+Nengo can run network simulations in several different "modes", which provide different trade-offs between speed and accuracy.
+"Default mode" performs spiking simulations in the manner described above.
+In "rate mode", the simulator does not integrate the membrane dynamic equations, but instead calculates an analytical spike rate for each neuron based on its net input.
+Rate mode typically runs slightly faster than default mode, but it ignores high-frequency spike-related fluctuations in synaptic current.
+Another particularly useful mode is "direct mode".
+In direct mode, neurons are not simulated at all.
+Rather, the abstract dynamic equations, \textbf{$\mathbf{\frac{\mathrm{d}x}{\mathrm{dt}}=}\mathbf{f}(\mathbf{x}(t))+\mathbf{u}(t)$}, are simulated directly.
+These simulations are very fast, but they omit both spike-related fluctuations and also any static biases that arise from approximation of idealized functions with sums of spike rates.
+(Such biases are often substantial, particularly as spike rates saturate outside a population's normal operating range.) Despite this high level of abstraction, direct mode is very useful for rapidly prototyping and debugging high-level network function, and also for exploring differences between idealized network function and its approximation by neurons.
+
+We recently designed an additional new simulation mode \cite{Tripp2015}, which we call "population mode".
+In population mode, the simulator runs efficient surrogate models of population output.
+These models are slightly more computationally intensive than direct mode, but they are more realistic in that they include both static biases and spike-related fluctuations (similar to default mode).
+Specifically, they model the difference between the output of each population in default mode and direct mode, i.e.
+$\mathbf{e}(\mathbf{x},t)=\mathbf{f}(\mathbf{x})-\sum_{i}a_{i}(\mathbf{x)}\mathbf{d}_{i}^{\mathbf{f}}$.
+This difference is modelled as $\mathbf{e}(\mathbf{x},t)=\mathbf{s}_{\mathbf{x}}(t) + \mathbf{b}({\mathbf{x})}$, where $\mathbf{s}_{\mathbf{x}}(t)$ is the spike-related fluctuations (the difference between default and rate modes), and $\mathbf{b}(\mathbf{x})$ is the static bias (the difference between rate and direct modes).
+The bias model $\bar{\mathbf{b}}(\mathbf{x}) \approx \mathbf{b}(\mathbf{x})$ consists of an interpolation or regression between sampled values, and $\bar{\mathbf{s}}_{\mathbf{x}} \approx \mathbf{s}_{\mathbf{x}}$ consists of Gaussian white noise that is filtered to approximate the spectrum of the population's default-mode output fluctuations.
+This approach has been demonstrated in models of neural integrators, nonlinear oscillators, and dynamic neural fields \cite{Tripp2015}.
+
+The system introduced here is an implementation of population mode for field-programmable gate arrays.
+It is also the first Nengo backend to support population mode, since the approach was developed in stand-alone software.
 
@@
 \section{Results}
-
+- ??? Schematics could come from the Nengo visualizer, for consistency and ease.
 % integrator (1d)
@@
 
+\section{Large-scale simulation}
+
+- Bryan??? Describe the V1 model, simulate it and show results... big as possible
+
 \section{Discussion}
@@
 
+\subsection{Conclusion}
+
+- Bryan???
+
 \bibliographystyle{acm}

### 2015-05-19 docs/paper/paper.tex
adding draft of V1 model section

@@
 
-- Bryan??? Describe the V1 model, simulate it and show results... big as possible
+As an example of a large, real-time simulation, we adapted an existing model of visual motion sensitivity (ref Aziz).
+The original version of this model illustated a mechanism by which motion selectivity might arise from recurrent connections within V1.
+This mechanism is interesting because recurrent connections are anatomically prominent, but absent from classical models of motion selectivity (ref Hubel).
+The recurrent connections are optimized, using the NEF, to approximate two-dimensional damped linear oscillators.
+Each oscillator receives subcortical burst input due to visual features at multiple spatial phases within the receptive field.
+Inputs from different parts of the receptive field drive the oscillator state in different directions.
+As an example, consider a population in which input from the left half of the receptive field drives the oscillator state to the right, and input from the right half drives the state to the left.
+Suppose the oscillator state is initially zero, and a visual feature travels across the receptive field from left to right.
+When the input is on the left, it will drive the oscillator state from zero to the right.
+This will begin a damped oscillation, and a short time later the state will be travelling to the left.
+If input from the right half of the receptive field arrives at this time, it will drive the oscillator state further to the left, amplifying the oscillation.
+Because the oscillator is damped, neurons that represent later states in the cycle are only active following this kind of constructive interference.
+These late-phase neurons provide output to MT populations.
+MT populations therefore respond to motion at speeds that maximize constructive interference in the corresponding V1 oscillators.
+
+Our implementation of the model included 4500 oscillator populations of 200 motion-selective neurons each.
+These neurons had small, Gabor-shaped spatial receptive fields (roughly 40-pixel diameter).
+The populations varied in the positions of their receptive fields, and their direction and speed preferences.
+Six additional populations modelled a small part of area MT.
+Each of the MT populations received converging input from 100 V1 populations that spanned a receptive fields 400 pixels in diameter.
+
+The model accepted 400x400 pixel, 100 frame/s video as input.
+As an abstraction of subcortical processing, each video frame was convolved with a difference-of-Gaussians filter.
+The V1 and MT populations ran at 1000 steps/s.
+Input to the neural model consisted of a 400x400-pixel input, $\mathbf{u}$, which was calculated from the most recent video frame.
+Specifically, pixels of $\mathbf{u}$ in which the filtered image both exceeded a threshold and exceeded its value in the previous frame were set to a high value for the next 6ms.
+This modelled bursting in magnocellular lateral geniculate nucleus (ref Aziz).
+
+Each V1 population received weighted input from $\mathbf{u}$ in a spatially restricted receptive field.
+To set the weights, the difference-of-Gaussian functions used in the subcortical model were combined in a weighted sum, and the weights were optimized so that the weighted sum approximated Gabor functions.
+Groups of three Gabor functions with the same orientation and spatial frequency, but different phases, defined the inputs to groups of six V1 populations with the same spatial receptive field but different direction and speed preferences (see Figure) [TODO: picture of left, centre, right Gabor functions in 400x400 image; picture of dots].
+The abstract population dynamics of the leftward-sensitive populations was
+
+\[
+\dot{\mathbf{x}} = 
+\left[
+\begin{array}{cc}
+-\zeta & \omega \\ 
+-\omega & -\zeta \\ 
+\end{array}
+\right]
+\mathbf{x} + 
+\left[
+\begin{array}{c}
+\mathbf{w}_l^{T} - \mathbf{w}_r^{T} \\ 
+-\mathbf{w}_c^{T} \\ 
+\end{array}
+\right]
+\mathbf{u}
+\]
+
+where $\omega$ is the oscillator frequency, $\zeta$ is the damping ratio, and $\mathbf{w}_l$, $\mathbf{w}_c$, and $\mathbf{w}_r$ are vectors of the weights for the left, centre, and right Gabors, respectively.
+The first state variable was therefore driven to increase in response to input on the left and decrease in response to input on the right of the receptive field.
+The second state variable was driven to decrease in response to input in the centre of the receptive field.
+A stimulus passing across the receptive field from left to right at the right speed would therefore interfere constructively with the (clockwise) intrinsic oscillation of the population state.
+Input to the leftward-sensitive populations was the same except that the sign of the left and right signs were exchanged.
+To limit the size of the model, we only included populations sensitive to leftward and rightward motion (Our system is not able to simulate all of primate V1 on a single FPGA.) Full source code is available from the authors' web site (ref).
+
+Figure XX shows simulations with coherent random dot stimuli.
+On the left are decoded outputs of representative V1 and MT populations from a spiking software simulation (default mode).
+On the right are the corresponding outputs from a population-mode simulation on the FPGA.
+The outputs are similar in each case.
+However, while the FPGA simulation ran in real time, the software software simulation ran at 1/???
+real-time and consumed much more power.
 

### 2015-05-19 v1/motion_inputs.py
cleaning up stimulus/input generation code

@@ -3,27 +3,7 @@ import numpy as np
 from scipy import ndimage as nd
 import cPickle as pickle
 
-def make_DOG(inner_sigma, x): 
-    """ Make a 1D difference-of-Gaussians kernel. """
-    outer_sigma = inner_sigma*4
-    inner_gaussian = 1./np.sqrt(2*np.pi)/inner_sigma * np.exp(-x**2/2./inner_sigma**2)
-    outer_gaussian = 1./np.sqrt(2*np.pi)/outer_sigma * np.exp(-x**2/2./outer_sigma**2)
-    return inner_gaussian - outer_gaussian
-
-def make_DOG2D(inner_sigma, x, y, centre):
-    """ Make a 2D difference-of-Gaussians kernel. """
-    outer_sigma = inner_sigma*4
-    inner_gaussian = 1./(2*np.pi*inner_sigma**2) * np.exp(-((x-centre[0])**2 + (y-centre[1])**2)/2/inner_sigma**2)
-    outer_gaussian = 1./(2*np.pi*outer_sigma**2) * np.exp(-((x-centre[0])**2 + (y-centre[1])**2)/2/outer_sigma**2)
-    return inner_gaussian - outer_gaussian
-
-def make_gabor(x, y, centre, frequency, phase, sigma, angle): 
-    """ Make a Gabor kernel. """
-    xx = x*np.cos(angle) - y*np.sin(angle)
-    yy = x*np.sin(angle) + y*np.cos(angle)
-    xx = xx-centre[0]
-    yy = yy-centre[1]
-    return np.cos(frequency*xx + phase) * np.exp(-xx**2/2./sigma[0]**2 -yy**2/2./sigma[1]**2)
+# Note: I've been running this in Python 2.7 -- not sure if PsychoPy works in Python 3
 
 #Assuming frame rate of 100Hz (will sample-and-hold for 1000Hz neural simulation)  
 def make_dots(n_frames, coherence=0.5, speed=0.1, direction=180):
@@ -70,129 +50,9 @@ def make_dots(n_frames, coherence=0.5, speed=0.1, direction=180):
         
     return dot_images
 
-def make_DOGs(inner_sigma, centres, x, y):
-    """ Make a grid of 2D difference-of-Gaussian kernels, flattened into vectors. """
-    DOGs = np.zeros((x.shape[0]*x.shape[1], len(centres)**2))
-    c = 0
-    for cx in centres: 
-        for cy in centres: 
[…]

### 2015-05-19 v1/v1.py

@@ -3,22 +3,29 @@ from scipy import ndimage as nd
 import matplotlib.pyplot as plt
 import pickle
 
-def make_DOG2(inner_sigma, x):
+def make_DOG(inner_sigma, x):
+    """
+    Creates a difference-of-Gaussians kernel.
+    
+    Arguments: 
+    ---------
+    inner_sigma: width of centre gaussian (surround width is a constant multiple of this)
+    x: domain in one dimension (should have 0 as a centre)  
+    """
     y = x
     outer_sigma = inner_sigma*5
     X, Y = np.meshgrid(x, y)
     inner_gaussian = 1./(2.*np.pi*inner_sigma) * np.exp(-(X**2 + Y**2)/2./inner_sigma**2) 
     outer_gaussian = 1./(2.*np.pi*outer_sigma) * np.exp(-(X**2 + Y**2)/2./outer_sigma**2) 
-#     return inner_gaussian - outer_gaussian
-    return inner_gaussian - outer_gaussian / 2
+    return inner_gaussian - outer_gaussian/2 #weaker surround works better with our weights, which don't account for bursts 
 
 def make_gabor(x, frequency, phase, sigma): 
-    """ Make a Gabor kernel. """
+    """ Makes a 1D Gabor cross-section. """
     return np.cos(frequency*x + phase) * np.exp(-x**2/2./sigma**2)
 
 def make_gabors(x):
     """
-    Creates 1D Gabor cross-sections 
+    Creates 1D Gabor cross-sections. 
     
     Returns:
     -------
@@ -37,10 +44,16 @@ def make_gabors(x):
     return gaborx, gabory
      
 def filter_dots(dots, DOG):
+    """
+    Convolves dot image with kernel. 
+    """
     return nd.filters.convolve(dots, DOG)
 
 
 def shift_kernel(kernel, shape, centre):
+    """
+    Moves kernel to different point in image (this is needed for finding weights). 
+    """
     h, w = kernel.shape
     assert(h % 2 == 1)
     assert(w % 2 == 1)
@@ -52,9 +65,8 @@ def shift_kernel(kernel, shape, centre):
     ind_h = centre[0] + np.arange(0, 2*half_h+1, dtype='int')    
     ind_w = centre[1] + np.arange(0, 2*half_w+1, dtype='int')
     result[ind_h[:,np.newaxis], ind_w] = kernel
[…]

### 2015-05-19 v1/v1.py
adding code for dot stimuli at multiple speeds

@@ -189,21 +189,21 @@ def check_gabor_approx():
     plt.show()
 
  
-def filter_default_dots():
+def filter_default_dots(dot_file):
     """
     Filter dot stimulus with DOG filter, where both dot stimulus and DOG params 
     are default values. 
     """
     x = np.arange(-40, 41, 1)
-    DOG2 = make_DOG(3, x)  
-    dots = pickle.load(open("/Users/bptripp/code/nengo-FPGA/v1/dot-images-coh1-2000ms-s02.p", "rb" ), encoding='latin1')  
+    DOG = make_DOG(3, x)  
+    dots = pickle.load(open(dot_file, "rb" ), encoding='latin1')  
     dots = dots - np.min(dots)
     filtered = np.zeros_like(dots)
     for i in range(dots.shape[2]):
         print('processing frame ' + str(i))    
-        filtered[:,:,i] = filter_dots(np.squeeze(dots[:,:,i]), DOG2)
+        filtered[:,:,i] = filter_dots(np.squeeze(dots[:,:,i]), DOG)
      
-    pickle.dump(filtered, open("/Users/bptripp/code/nengo-FPGA/v1/dog-filtered-dots-new.p", "wb" ))
+    pickle.dump(filtered, open(dot_file.replace('.p', '-filt.p'), "wb" ))
     
 def pure_gabor():
     """
@@ -237,7 +237,7 @@ def pure_gabor():
                 result[j,k,i] = np.sum(gabor * region)
     return result
 
-def get_default_inputs():
+def make_default_inputs(filtered_dot_file):
     """
     Obtain inputs for the neural model using default values for dot stimulus, DOG, 
     etc. 
@@ -249,24 +249,26 @@ def get_default_inputs():
 #     w_kernel = pickle.load(open("/Users/bptripp/code/nengo-FPGA/v1/dog-gabor-weights-3.p", "rb" ))
 #     w_kernel = pickle.load(open("/Users/bptripp/code/nengo-FPGA/v1/dog-gabor-weights-2.p", "rb" ))
 #     w_kernel = pickle.load(open("/Users/bptripp/code/nengo-FPGA/v1/dog-gabor-weights-4.p", "rb" ))
-#     filtered = pickle.load(open("/Users/bptripp/code/nengo-FPGA/v1/dog-filtered-dots.p", "rb" ))
[…]

### 2015-05-21 docs/paper/paper.tex
adding conclusion to paper

@@
 \subsection{Conclusion}
-
-- Bryan???
+We have presented a new neuromorphic system based on commodity FPGA hardware.
+We emphasize that FPGA expertise is not needed to use this system.
+Models can be developed for the system using the Nengo simulator, and implemented in hardware with a single additional Python command.
+Compared to existing FPGA approaches to neural simulation, our use of surrogate population models allows a substantial increase in model scale, with associated constraints (e.g.
+on fan-in, fan-out, and population dimension), and loss of fidelity.
+However, many existing models are already consistent with these constraints, and further loss of fidelity relative to real neural systems may not be much greater than with point-neuron models, which are already highly simplified.
+The increase in scale raises the possibility of simulating more extensive neural systems that are capable of richer interactions with the physical world. 
+Because population inputs are transport-agnostic, this system could also potentially be used for very large, power-efficient simulations, by connecting multiple FGPAs through Ethernet or the FPGAs' general-purpose input/output pins.
+In summary, this work provides a practical and accessible new way to implement neuromorphic systems.
 

### 2015-05-21 v1/v1.py
fixing MT receptive field (via inputs); reinstating constant damping ratio

@@ -255,18 +255,20 @@ def make_default_inputs(filtered_dot_file):
     n_inputs = 200
     centres = np.random.randn(n_inputs,2)
     mag = np.sum(centres**2, axis=1)**(1/2)
-    centres = 180 * centres.T / mag
+    centres = 155 * centres.T / mag
     centres = centres * np.random.rand(n_inputs)**(1/2)
     centres = np.round(centres).astype('int')
-
+    centres = centres + 199 #put in image centre
+    
     inputs = get_inputs(w_kernel, filtered, centres.T)
+    
     pickle.dump(inputs, open(filtered_dot_file.replace('-filt.p', '-inputs.p'), "wb" ))
 
 
 # dot_file = "/Users/bptripp/code/nengo-FPGA/v1/dot-images-coh1-s04.p"
 # filter_default_dots(dot_file)
 
-make_default_inputs("/Users/bptripp/code/nengo-FPGA/v1/dot-images-coh1-s005-filt.p")
+make_default_inputs("/Users/bptripp/code/nengo-FPGA/v1/dot-images-coh1-s04-filt.p")
 
 # check_gabor_approx()
 # get_default_weights()

### 2015-09-28 docs/paper/principal-components.tex
adding figures, text, code on shared principal components

\documentclass[]{article}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[outdir=./]{epstopdf}
\begin{document}

\section{Principal component sharing}
%\maketitle

Tripp (2015) approximated population outputs by combining surrogate models of time-averaged output $\hat{\mathbf{f}}(\mathbf{x})$ and spike-related fluctuations $\mathbf{n}(t)$.

In the present study we use the same noise term, but we model the static output differently. Specifically, we approximate the output as a weighted sum of principal components $\mathbf{p}_k(\mathbf{x})$ of the population's tuning curves, i.e.
\begin{equation}
\hat{\mathbf{f}}(\mathbf{x}) =  \sum_{k=1}^m \phi_k \mathbf{p}_k(\mathbf{x})
\end{equation}
where $m$ is the number of principal components used in the approximation.

Whereas Tripp (2015) used interpolation to model each population output $\hat{\mathbf{f}}(\mathbf{x})$, here we use interpolation to model $\mathbf{p}_k(\mathbf{x})$, and then combine these linearly into approximations of the population outputs. This converges to the same result when many principal components are used, but it  
decouples the number of interpolations from the number of outputs. This is an advantage when a population has more distinct outputs than useful principal components.

The principal components of different populations that are drawn from the same parameters are typically quite similar, as illustrated in Figure \ref{fig:pc3}. The components are usually moderately consistent between smaller populations (left panel of Figure \ref{fig:pc3}) and quite consistent between large populations (right panel of Figure \ref{fig:pc3}). In light of this consistency, we do not use the unique principal components of each individual population, but instead we use ``shared" principal components that approximate those of all the populations in a population unit. This reduces the required number of interpolation tables by up to three orders of magnitude, and allows us to leave the interpolation tables in on-chip memory throughout a simulation. To simulate each distinct population in a population unit, we need only load the unique mixing weights $\phi$ for that population (in addition to the noise parameters). This efficiency at the population-unit level is our main motivation for interpolating principal components instead of outputs.

It turns out that principal components of different populations can also be quite similar, even if there are substantial differences in population parameters. This is illustrated in Figure \ref{fig:pc-examples}. For this reason we do not require that all populations in a population unit have exactly the same parameters. Instead, we cluster populations according to principal component similarity and assign each cluster to a population unit.

To find mixing weights $\phi_k$ for the principal components, we first find unregularized optimal decoders $d_i^k$ for the $i^{th}$ neuron and the $k^{th}$ principal component. This allows us to estimate the variance of the noise associated with the $k^{th}$ principal component as $\sigma_k^2 = \sum_i (d_i^k)^2 \sigma^2$, where $\sigma^2$ is the noise variance of each neuron (we typically assume this to be uniform). Finally, we find regularized optimal decoders of the principal components by choosing $\phi_k$ to minimize
\begin{equation}
\int_{\mathbf{x}} d\mathbf{x} \left( \hat{\mathbf{f}}(\mathbf{x}) -  \sum_{k=1}^m \phi_k \mathbf{p}_k(\mathbf{x}) \right)^2 + \sum_{k=1}^m \phi_k \sigma_k^2
\end{equation}

To share principal components across a population unit, we use the same $\sigma_k^2$, and simply replace the population principal components with the shared principal components.

\begin{figure}
\centering{}
\includegraphics[scale=0.4]{PC3.eps}
\caption{The third principal components of tuning curves of random populations drawn from the same parameter distribution (parameter set \#6). The left panel shows the third principal components of ten random draws of 50 neurons. The right shows the third principal components of ten random draws of 1000 neurons. The principal components are similar in each case, and very similar across the large populations. The third principal component is shown as an example; other components have similar behaviour.}
\label{fig:pc3}
\end{figure}

\begin{figure}
\centering{}
\includegraphics[scale=0.4]{PC-examples.eps}
\caption{Tuning curves of four different example populations and their principal components. Each column corresponds to a population with different parameters (parameter sets 1, 6, 7, and 10). The top panels are 50 tuning curves drawn from populations of 1000 neurons. The bottom panels are the first six principal components of each population's tuning curves. Despite differences in parameters between the three leftmost populations, their principal components are nearly the same or nearly opposite. (To facilitate comparison across populations, the negatives of some principal components are plotted where these are more similar to corresponding components in the leftmost column.) As a counter-example, the population on the right has a very different distribution of intercepts, and very different principal components as well. }
\label{fig:pc-examples}
\end{figure}

Figure \ref{fig:shared-pcs} shows examples of static error associated with these various approximation schemes. The solid dark traces are static errors in function approximation directly from neuron tuning curves. The (fairly similar) solid light lines are static errors in function approximation from principal components of the same population.

The dashed lines are approximation from principal components of other populations, to illustrate the potential loss in simulation accuracy associated with using shared principal components. The darker dashed line is from principal components of a larger population with the same parameters, and the lighter dashed line is from principal components of a population of the same size but different parameters. Both of these differences result in quite different static error functions than the decoding directly from tuning curves. Importantly, the error still has approximately the right frequency content, and the magnitude varies appropriately depending on the decoded function (left vs. right panels). However, the differences illustrate the importance of clustering populations so that shared principal components are as similar as possible to each population's individual principal components.

\begin{figure}
\centering{}
\includegraphics[scale=0.4]{shared-PCs.eps}
\caption{Comparison of bias error with different approximation schemes. The left panel shows bias errors in decoding the function $\sin(\pi x)$. The different error plots are offset vertically for clarity; the scale bar indicates an error magnitude of 0.1. The solid dark line shows decoding error from a population of 100 neurons (parameter set \#6). The solid light line below it shows decoding error from the first 15 principal components of the same population. The dark dashed line shows decoding error from the first 15 principal components of a larger population (500 neurons) with the same parameter distribution. The principal components of the larger population are smoother, and this error function has lower high-frequency content. Finally, the light dashed line shows decoding error from the first 15 principal components of a 100-neuron population with different parameters (parameter set \#1). The right panel shows analogous errors for decoding of $\sin(3 \pi x)$. As is typical, this higher-frequency function is not decoded as well, because it requires greater involvement of minor (noisier) principal components. } 
\label{fig:shared-pcs}
\end{figure}

\end{document}

### 2015-09-28 ml/expPCs.m

@@ -1,0 +1,147 @@
+% exploring variations in principal components across populations with
+% different parameters and sizes
+
+
+points = -1:.01:1;
+
+%%
+rng(1)
+figure 
+
+n = [50 1000];
+reps = 10;
+
+repPCs = zeros(length(points), reps);
+for i = 1:length(n)
+    subplot(1,length(n), i)
+    for j = 1:reps
+        sg = getPMSpikeGenerators(n(i));
+
+        pop = CosinePopulation(1, sg{6}, 'test');
+        rates = getRates(pop, points, 0, 0);
+
+        [U, S, V] = svd(rates);
+
+        ind = 3;
+        if V(1,ind) > V(100,ind)
+            repPCs(:,j) = V(:,ind);
+        else 
+            repPCs(:,j) = -V(:,ind);
+        end
+    end
+    plot(points, repPCs, 'k')
+    set(gca, 'FontSize', 18)
+    xlabel('$x$', 'FontSize', 24, 'Interpreter', 'LaTex')
+    set(gca, 'YLim', [-.12 .17])
+end
+set(gcf, 'Position', [440 479 690 319])
+
+%%
+rng(1)
+sg = getPMSpikeGenerators(1000);
+sgInd = [1 6 7 10];
+npc = 6;
+figure
+for i = 1:length(sgInd)
+    pop = CosinePopulation(1, sg{sgInd(i)}, 'test');
+    rates = getRates(pop, points, 0, 0);
+    [U, S, V] = svd(rates);
+    
+    PCs = V(:,1:npc);
+    if i == 1 
+        firstPCs = PCs;
+    else % change signs of PCs as needed to match first population
+        prod = PCs' * firstPCs;
+        for j = 1:npc
+            if prod(j,j) < 0
+                PCs(:,j) = -PCs(:,j);
+            end
+        end
+    end
+    
+    subplot(2, length(sgInd), i)
+    plot(points, rates(1:50,:), 'k')
+    set(gca, 'XTick', [])   
+    if i == 1, ylabel('Spike Rate (spikes/s)', 'FontSize', 18), end
+    set(gca, 'FontSize', 18)
+    
+    subplot(2, length(sgInd), length(sgInd)+i)
+    plot(points, PCs, 'k')
+    set(gca, 'YLim', [-.18 .18])
+    if i == 1, ylabel('PCs', 'FontSize', 18), end
[…]

### 2015-09-28 ml/getPMSpikeGenerators.m

@@ -1,0 +1,68 @@
+function sg = getPMSpikeGenerators(n, varargin)
+    % sg = getPMSpikeGenerators(n) creates a selection of spike generators 
+    % with a wide range of parameters. 
+    % 
+    % n: number of neurons to include in each spike generator
+    % sg: cell array of spike generators
+    
+    if nargin > 1
+        rng(varargin{1});
+    end
+
+    %parameters of populations with gaussian-distributed intercepts 
+    gp = [ 
+        .005 .04 2/3 2 2; 
+        .002 .02 2/3 1 .5; 
+        .002 .02 1 2 1; 
+        .003 .03 1.5 3 .2; 
+        .004 .01 1 3 .2; 
+        ]; 
+    
+    % parameters of populations with uniform distributions 
+    up = [
+        .002 .02 -1 1 200 400; %defalt parameters for years
+        .003 .02 -.9 .9 50 100; 
+        .005 .02 -1 1 30 80; 
+        .002 .1  -1 1 50 100; 
+        .002 .1  0 1 50 100;
+        ];
+    
+    sg = cell(size(gp,1)+size(gp,2), 1);
+    for i = 1:size(gp,1)
+        sg{i} = getGaussianSG(n, gp(i,1), gp(i,2), gp(i,3), gp(i,4), gp(i,5));
+    end
+    
+    for i = 1:size(up,1)
+        sg{size(gp,1)+i} = getUniformSD(n, up(i,1), up(i,2), up(i,3), up(i,4), up(i,5), up(i,6));
+    end
+    
+%     for i = 1:length(sg)
+%         plotPopulation(sg{i}, 2), pause
+%     end
+
+end
+
+function sg = getGaussianSG(n, tauRef, tauRC, interceptSD, scaleShape, scaleScale)
+    intercepts = interceptSD * randn(1,n);
+    maxRates = 50+50*rand(1,n); %dummy values
+    V0 = 0;
+    sg = LIFSpikeGenerator(.0005, tauRef, tauRC, intercepts, maxRates, V0);
+    sg.scales = gamrnd(scaleShape, scaleScale*ones(1,n));
+    sg.biases = 1 - sg.scales .* intercepts;
+end
+
+function sg = getUniformSD(n, tauRef, tauRC, interceptLow, interceptHigh, maxRateLow, maxRateHigh)
+    intercepts = interceptLow + (interceptHigh-interceptLow)*rand(1,n);
[…]
